Blocking Toxin Function and Modulating the Gut Microbiota: Caffeic Acid and its Derivatives as Potential Treatments for Clostridioides difficile Infection

Version 1

Abstract

Clostridioides difficile infection (CDI) is the leading cause of hospital- acquired diarrhea that seriously threatens public health. The disruption of normal gut microbiota by the use of broad-spectrum antimicrobial agents enables C. difficile to proliferate in the colon. The emergence and prevalence of hypervirulent C. difficile strains result in increased morbidity, mortality, and recurrence rates of CDI, thus creating a pressing need for novel therapeutics. The multi-domain toxins TcdA and TcdB are the primary determinants of CDI pathogenesis, rendering them ideal drug targets in the anti-virulence paradigm. In this study, we identified caffeic acid and its derivatives as active inhibitors of TcdB via a cell-based high-throughput phenotypic screening. Further mechanistic investigations revealed that caffeic acid phenethyl ester (CAPE) could directly bind to TcdB, thus suppressing InsP6-induced autoproteolysis and inhibiting glucosyltransferase activity. CAPE treatment remarkably reduces the pathology of CDI in a murine infection model in terms of alleviated diarrhea symptoms, decreased bacterial colonization, and relieved histopathological lesions. Moreover, CAPE treatment of C. difficile-challenged mice induces a remarkable increase in the diversity and composition of the gut microbiota (e.g., Bacteroides) and alterations of gut metabolites (e.g., adenosine, D-proline, and melatonin), which might partially contribute to the therapeutic outcomes of CAPE against CDI. Our results reveal the potential of CAPE as a therapeutic for the management of CDI, or CAPE might serve as a lead compound for the development of antivirulence drugs targeting TcdB.

Introduction

Clostridioides difficile (C. difficile) is an obligate anaerobic, spore-forming, gram-positive bacterial pathogen that is the most common cause of hospital- acquired and antibiotic-associated gastrointestinal diseases, including mild diarrhea and life-threatening pseudomembranous colitis 1, 2, 3. In the United States, C. difficile is responsible for almost half a million infections and approximately 29,000 deaths per year, resulting in an estimated annual healthcare-associated cost of up to $4.8 billion 4, 5. Importantly, the incidence, severity, mortality and healthcare costs of C. difficile infection (CDI) have dramatically increased in the 21st century, at least partially owing to the frequent use of antimicrobial agents and the emergence of hypervirulent strains (e.g., PCR ribotypes 027 and 078) 6, 7, 8. Limited options are available for the treatment of CDI, while antibiotics including metronidazole, fidaxomicin and vancomycin, are the mainstays of clinical practice 9, 10. However, antibiotic treatment is ineffective in more than 35% of CDI patients and can lead infection recurrence 11, 12. Moreover, the rapid evolution of antibiotic resistance in C. difficile has further compromised the therapeutic efficacy of current treatment options 13. Therefore, the high morbidity and mortality, high recurrence rates, and emergence of multiple antibiotic-resistant isolates have made this pathogen a major threat to public health; these factors highlight the urgent need for novel therapies.

C. difficile widely exists in the environment and in the intestinal tract of humans and animals and is mainly transmitted via the fecal-oral route 14. The pathogenesis of CDI is correlated with the use of antibiotics, which impair the structure and composition of the normal gut microbiota, thus allowing C. difficile to proliferate in the colon 14. Therefore, previous antibiotic therapy is the major risk factor for the development of CDI. In the colon, the dormant spores of C. difficile germinate into oxygen-sensitive vegetative cells that produce toxins and cause symptoms of disease 15. C. difficile secretes two large clostridial exotoxins (TcdA and TcdB), which are directly responsible for the clinical pathology of CDI 16, 17. Epidemiological studies suggest that most disease- causing C. difficile strains have genes that encode TcdB 18, 19. TcdA and TcdB have identical functional domains: a N-terminal glucosyltransferase catalytic domain (GTD), a cysteine protease active domain (CPD), a transmembrane domain (DRBD), and a C-terminal receptor binding domain (CROPs) 20. When secreted into the colon, both toxins are internalized by host cells through receptor-mediated, clathrin-dependent endocytosis. Following a decrease in endosomal pH, the toxins undergo conformational changes in the DRBD, resulting in insertion of the toxin into the endosome membrane, pore formation, and the release of the GTD and CPD domains into the cytosol. Then, CPD binding with cytosolic inositol hexakisphosphate (InsP6) activates its protease activity, leading to the cleavage and release of the GTD. Eventually, the GTD utilizes UDP-glucose to glucosylate Rho and/or Ras family GTPases, thus disrupting GTPase signaling and resulting in deleterious effects, including cytopathic “rounding”, dysregulation of the actin cytoskeleton, apoptosis, and pyroptosis 1. Although both toxins are cytotoxic to host cells, TcdB shows greater toxicity than TcdA 21.

Disruption of the gut microbiota by antimicrobial agents results in high recurrence rates, which limits the application of antimicrobial therapy for CDI 14. Consequently, there is a critical need to develop nonantibiotic treatments for CDI, either as standalone or augmented therapies. Importantly, strategies targeting bacterial virulence factors, including toxins, are termed anti-virulence strategies and have attracted interest in the 21st century 22. Owing to its essential role in C. difficile pathogenesis, TcdB has become an ideal target for the development of anti-virulence drugs 23. In this study, caffeic acid and a series of its derivatives were found to inhibit TcdB-mediated cell rounding via cell-based screening. Among these potential inhibitors, caffeic acid phenethyl ester (CAPE) had the greatest inhibitory effect. Mechanistically, CAPE bound to both full-length TcdB and the GTD fragment, leading to alterations in the secondary structure. This interaction suppressed the InsP6-mediated self- cleavage of TcdB, as well as the enzymatic activity of GTD, thereby decreasing the level of glucosylated Rac1 in TcdB-treated host cells and further inhibiting the cytopathology and cytotoxicity of TcdB. Oral administration of CAPE significantly protected against CDI in models of animal infection. In addition to direct toxin inhibition, the therapeutic effect may also be attributed to CAPE- induced alteration of the gut microbiota (e.g., Bacteroides) as well as gut metabolites (e.g., adenosine, D-proline, and melatonin). Overall, our study revealed that caffeic acid and its derivatives, particularly CAPE, may serve as promising lead compounds for the development of anti-virulence drugs to treat CDI.

Results

CAPE inhibits cell rounding induced by TcdB

TcdB can inactivate Rho GTPases via glucosylation, leading to the loss of cytoskeletal structure and cell-rounding. Consequently, we utilized this phenotype to identify potential TcdB inhibitors from 2076 chemicals, including 1515 FDA-approved drugs and 561 natural compounds. The minimal concentration of recombinant His6-TcdB (0.2 ng/mL) that induced 100% cell rounding within 1 h was used for drug screening. This concentration of TcdB was preincubated with 8 μg/mL of the individual compounds for 1 h before the compounds were added to Vero cell cultures. At 1 h post-TcdB treatment, cell morphology was visualized under a light microscope (Figure 1A). These analyses allowed us to identify 22 candidate compounds with potent inhibitory effects (Figure 1B). Interestingly, six out of the 22 compounds were caffeic acid derivatives. Caffeic acid itself only slightly inhibited TcdB-induced cell rounding, with a 50% inhibitory concentration (IC50) of 527.5 μg/mL (Table S1). Notably, except for rosmarinic acid, the IC50 values for the other screened caffeic acid derivatives were less than 6 μg/mL, suggesting that these compounds potently inhibited TcdB-mediated toxicity (Table S1). CAPE is the phenethyl alcohol ester of caffeic acid and a main active component of propolis with broad pharmacological properties, such as anticancer, anti-inflammatory, antiviral, antibacterial, antifungal, antioxidant, and immunomodulatory activities 24. As CAPE has the lowest IC50 (3.0 μg/mL) among the active caffeic acid derivatives, it was selected as the representative compound for subsequent investigations (Table S1). After preincubation with increasing concentrations of CAPE, dose- dependent inhibition of cell rounding was observed in both Vero and HeLa cells (Figure 1C). In addition, we also determined whether inhibition of TcdB by CAPE could lead to loss of toxin function in vivo using a murine systemic intoxication model. A lethal dose of recombinant His6-TcdB (30 ng) was preincubated with either DMSO or CAPE (2 or 4 μg/mL) for 1 h at 37 °C, followed by intraperitoneal injection into mice. Notably, all mice in the DMSO group (n=10) died within 36 h after toxin injection. However, 100% of mice pretreated with CAPE 4 (μg/mL) toxin survived at the monitoring endpoint (72 h) (n=10) (Figure 1D). Taken together, these data indicate that CAPE is a potent toxin inhibitor of TcdB.

CAPE suppresses TcdB-mediated cell rounding. (A) Schematic of the process used for the identification of small molecules that inhibit the cell rounding phenotype induced by TcdB. TcdB (0.2 ng/mL) was preincubated with 8 μg/mL of individual compounds prior to addition to the cell cultures. (B) Active compounds were identified from the chemical library consisting of 2076 chemicals. The red dots represent caffeic acid and its derivatives. (C) Dose-dependent CAPE inhibition of TcdB-induced rounding of Vero (upper) and HeLa (lower) cells. Scale bar, 100 μm. The dose titration curves are presented on the right. (D) The survival rates of mice that were intraperitoneally injected with CAPE-pretreated or native TcdB (30 ng). * p <0.05 and ** p <0.01 by log-rank (Mantel‒Cox) test.

CAPE decreases InsP6-induced autoproteolysis of TcdB and inhibits the enzymatic activity of GTD

TcdB is a multi-domain containing toxin that induces toxicity in host cells via an intricate and multistep mechanism. This mechanism involves (i) receptor- mediated endocytosis, (ii) pore formation and subsequent membrane translocation, (iii) autoproteolysis and release of GTD, and (iv) the glucosylation of small GTPases (Figure 2A). To determine the mechanism by which CAPE suppresses the cytotoxicity of TcdB, we first labeled full-length TcdB with Alexa Fluor 488; then, the concentration of the endocytosed toxin was determined via flow cytometry. Apparently, incubation of CAPE with TcdB did not significantly affect its entry into host cells (Figure 2B). Next, we stained CAPE-treated cells with LysoTracker, a fluorescent dye that accumulates in the acidic organelles of live cells. In contrast to the vacuolar-type H+-ATPase (V-ATPase) inhibitor bafilomycin A1 (Baf-A1), CAPE treatment did not alter the intensity of LysoTracker fluorescence staining in either THP-1 or Caco-2 cells, suggesting that inactivation of TcdB by CAPE is not due to changes in the pH of endosomal compartments (Figure 2C-D and Figure S1A-B).

Mechanism by which CAPE suppresses TcdB intoxication. (A)

Schematic illustration of the multistep mechanism of TcdB intoxication of host cells. (B) Determination of internalized TcdB levels. Alexa Fluor 488- labeled TcdB was incubated with 32 μg/mL CAPE for 30 min before it was added to the cell culture. The cellular uptake of TcdB was measured by flow cytometry analysis. (C) Lysosomal activity in host cells treated with CAPE was measured via LysoTracker staining. The nuclei were stained with DAPI. Bar, 100 μm. (D) The intensities of LysoTracker staining were quantified using ImageJ. * p <0.05; ** p <0.01 using one-way ANOVA. (E-F) CAPE inhibits InsP6- induced autoprocessing of full-length TcdB. Two micrograms of TcdB was preincubated with increasing concentrations of CAPE for 1 h. Self-cleavage of TcdB was initiated by the addition of 3 μg of InsP6 and the cells were incubated for 6 h at 37 °C. Reactions were terminated with 5x SDS loading buffer, and the release of GTD was visualized by CBB staining of the SDS‒PAGE gel (E). The level of released GTD in panel E was quantified by ImageJ (F) * p <0.05; ** p <0.01 using one-way ANOVA. (G) Dose-dependent inhibition of the GTD activity of TcdB by CAPE. Recombinant GTD of TcdB was mixed with different concentrations of CAPE (0-32 μg/mL) and incubated for 15 min at 25 °C. GTD activity was measured by the UDP-Glo™ glycosyltransferase assay. (H-I) Intracellular Rac1 glucosylation in cells treated with TcdB. Native or CAPE- pretreated TcdB was added to the cell culture and cells were incubated for 2 h. The cells were lysed, and the glucosylation of Rac1 was analyzed by western blotting using antibodies recognizing total Rac1 and unglycosylated Rac1 (H). The level of Rac1 glucosylation shown in H was quantified by ImageJ (I). * p <0.05; ** p <0.01 using one-way ANOVA.

These observations indicate that CAPE might exert its effects after membrane translocation. Indeed, preincubation of full-length TcdB with CAPE blocked the InsP6-induced autoproteolysis of TcdB, as revealed by the gradual decrease in the production of released GTP, followed by the addition of increasing amounts of CAPE in vitro (Figure 2E-F). Furthermore, we determined the glucosyltransferase activity using recombinant GTD in the absence or presence of different doses of CAPE. Strikingly, incubation with 2 μg/mL CAPE resulted in a significant reduction in GTD activity, whereas 32 μg/mL CAPE compromised 87.5% of the enzymatic activity (Figure 2G). Importantly, in response to suppression of InsP6-induced autoproteolysis and GTD function, treatment of TcdB-challenged cells with CAPE markedly decreased the level of glucosylated Rac1 in a dose-dependent manner (Figure 2H-I). Taken together, these results suggest that CAPE disrupts InsP6-induced autoproteolysis of TcdB and inhibits the enzymatic activity of GTD, thereby decreasing the glucosylation of Rac1 and alleviating cell rounding.

CAPE alters the secondary structure of full-length TcdB and GTD

To further clarify the mechanisms used by CAPE to inactivate TcdB, we determined the secondary structures of full-length TcdB and GTD in the absence or presence of CAPE by circular dichroism spectroscopy. Apparently, incubation of CAPE with full-length TcdB resulted in a significant increase in the α-helix content (15.4% to 21.0%) and a concomitant decrease in the β-pleated sheet content (34.5% to 20.9%) (Figure 3A-B). Similarly, the secondary structure of GTD was markedly altered upon incubation with CAPE, with the proportion of α-helices decreasing from 37.9% to 0.0% and the proportion of β- pleated sheets decreasing from 37.2% to 10.9% (Figure 3C-D). Taken together, these data suggest that CAPE may interact with full-length TcdB and GTD to induce conformational changes, thus impairing their functions.

CAPE alters the secondary structure of TcdB and GTD. (A-B)

The secondary structure of TcdB (A) or GTD (B) in the presence or absence of 32 μg/mL CAPE was determined by CD spectroscopy. (C-D) Calculated CD spectra of CAPE-treated TcdB (C) and GTD (D). The wavelength for CD spectroscopy was set at 190-250 nm.

CAPE directly interacts with full-length TcdB and GTD

To investigate the potential interaction between CAPE and TcdB or GTD, the binding constants were determined by surface plasmon resonance (SPR) analysis. The results showed that CAPE has strong binding affinity with full- length TcdB and GTD, and the KD values were 1.269×10-5 M and 1.007×10-3 M, respectively (Figure 4A-B). These data demonstrate the direct interaction between CAPE and full-length TcdB/GTD.

Direct interaction of CAPE with TcdB and GTD. (A-B)

Full-length TcdB or GTD was immobilized on a CM5 chip, followed by the injection of various concentrations of CAPE. The response unit (RU) value and affinities of CAPE interactions with TcdB (A) and GTD (B) were determined by a Biacore T100.

Next, we performed molecular docking and molecular simulation dynamics analyses to investigate the interaction between CAPE and GTD (Figure 5A). The results revealed that Leu265, Asp286, Thr465, Ile466, Leu519, and Trp520 in GTD were potential binding sites for CAPE (Figure 5A). The root mean square deviation (RMSD) of the GTD/CAPE complex was stable over a 100 ns period of simulation (Figure 5B). Further energy decomposition analysis showed that the side chains of Leu265, Thr465, Leu519, and Trp520 of GTD contributed to binding with CAPE via van der Waals interactions, among which Trp520 had the highest binding energy (Figure 5C). Moreover, we mutated each of the residues that were involved in binding and subsequently determined their glucosyltransferase activities in the presence or absence of CAPE. The enzymatic activities of these mutants were similar to that of wild-type GTD (Figure 5D). Mutations in GTDL265A, GTDT465A, GTDI466A, and GTDL519A did not affect the inhibitory effect of CAPE (Figure 5E-I). However, the inhibition of GTDW520A and GTDD286A by CAPE was notably compromised (Figure 5J-K).

Asp286 and Trp520 in the GTD are critical for CAPE-mediated inhibition of GTD. (A)

Three-dimensional structure determination of GTD with the CAPE complex was performed by molecular modeling. (B) The RMSD values of the GTD- CAPE complex. (C) Decomposition of the binding energy on a per-residue basis in the binding sites of the GTD-CAPE complex. (D) Glucosyltransferase activity of wild-type GTD and the GTD mutants. (E-K) Residual activity of wild-type GTD (E) and its mutants GTDL265A (F), GTDT465A (G), GTDI466A (H), GTDL519A (I), GTDW520A (J) and GTDD286A (K) in the presence of increasing concentrations of CAPE (0-4 μg/mL). * p <0.05; ** p <0.01 using one-way ANOVA.

CAPE treatment reduces the pathology of CDI in a murine model

The abovementioned observation that CAPE suppresses the function of TcdB prompted us to further evaluate the potential therapeutic effects of this compound in CDI. To this end, we employed a well-established murine model in which infected mice can develop symptoms resembling those observed in human CDIs (25). First, cytotoxicity measurements revealed that the viability of both THP-1 and Vero cells was not affected by CAPE treatment (Figure S2A- B), thus indicating low toxicity. Mice were preconditioned with drinking water containing 0.5 g/L cefoperazone for 5 days, followed by normal drinking water for 2 days. After challenge with C. difficile BAA-1870 via gavage, mice were then treated daily with either vehicle or 30 mg/kg CAPE by gavage throughout the disease course (day 0 to day 5) (Figure 6A). Severe weight loss in C. difficile-infected mice was observed on days 2 and 3 post infection and was accompanied by diarrhea (Figure 6B). Notably, compared to the infection control group, CAPE treatment significantly increased the dry/wet weight ratio of the mouse feces, suggesting the alleviation of diarrhea symptoms (Figure 6C). Correspondingly, the weight loss of mice was relieved upon receiving CAPE (Figure 6B). Moreover, fecal pellets collected on day 3 were used to determine the bacterial burden. The results revealed that CAPE treatment significantly decreased the number of C. difficile CFUs (Figure 6D). On day 3 post-C. difficile infection, the colon was collected for histopathological analysis. Remarkably, C. difficile-infected mice exhibited conspicuous colon edema, epithelial cell injury and inflammatory factor infiltration (Figure 6E). In contrast, these histopathological lesions were significantly less common in mice administered CAPE (Figure 6F). Taken together, these data demonstrate the efficacy of CAPE in the treatment of CDI.

CAPE treatment blocks the pathology of CDI in mice. (A)

Schematic of the process used for the establishment of the CDI mouse model. (B) Weight changes in CDI mice after CAPE or vehicle treatment (n=10). (C) The fecal dry/wet weight ratio in CDI mice treated with CAPE or vehicle (n=6). * p <0.05; ** p <0.01 using one-way ANOVA. (D) Bacterial colonization in the feces of CDI mice treated with CAPE or vehicle (n=15). * p <0.05; ** p <0.01 using the Mann‒Whitney U test. (E) Pathological analysis of colon sections from CDI mice treated with CAPE or vehicle. The black boxes are the regions that were magnified and are shown underneath. (F) Histopathological scores of vehicle-treated and CAPE-treated mice in terms of cell infiltration (orange), epithelial damage (purple), and edema (green). * p <0.05; ** p <0.01 using one-way ANOVA.

CAPE induces distinct microbiota changes in CAPE mice when compared with model mice

Antibiotic-induced dysbiosis of the gut microbiota promotes the development of CDI 25. Therefore, we next determined whether CAPE has any impact on gut microbial communities that may contribute to its therapeutic efficacy. To this end, on the 3rd day after C. difficile challenge, fecal samples from the model group and CAPE treatment group were collected and subjected to microbiota analysis via 16S rRNA gene sequencing. Remarkably, as reflected by the Venn diagram, CAPE treatment resulted in significant changes in the microbial composition compared to those in the model group, and only 16.42% of the operational taxonomic units (OTUs) were shared (Figure 7A). In addition, both the α- and β-diversity of the gut microbiota in CAPE-treated mice were significantly greater than those in model mice. Specifically, the increase in the Chao1 richness estimator of mice in the CAPE treatment group suggested a greater number of microbial species and increased evenness in the fecal samples (Figure 7B-C); the increase in the Shannon‒Wiener diversity index and the Simpson diversity index reflected a significant increase in microbiota diversity (Figure 7D-E). Principal coordinate analysis (PCoA) revealed distinct confidence ellipses between the bacterial communities of the CAPE-treated group and the model group, thereby revealing substantial differences in their microbiota profiles (Figure 7F).

The influence of CAPE treatment on the diversity and composition of the gut microbiota in CDI mice. (A)

Venn diagram illustrating the numbers of differentially abundant bacteria identified using a single model and their cooperation. (B-E) Quantification of the α diversity of the gut microbiota according to the observed species richness (B-C), Shannon index (D), and Simpson index (E). (F) PCoA showing β diversity based on the weighted UniFrac distance. (G-H) Relative abundance of the top 10 predominant bacteria classified at the phylum (G) and genus (H) levels. (I) Significant changes in flora among the model and CAPE groups, as measured by LEfSe analysis (LDA score [log 10] > 3.5). (J-M) Relative abundances of Bacteroides (J), Muribaculum (K), Faecalibaculum (L), and Parasutterella (M) in mice in the model and CAPE treatment groups. * p <0.05; ** p <0.01 using an unpaired Student’s t test.

A comparative analysis was performed to examine changes at the phylum level following CAPE treatment. Bacteroidota, Firmicutes, and Proteobacteria were the major constituents in both the CAPE treatment and model groups, suggesting that the administration of CAPE did not significantly alter the mouse microbiota at the phylum level (Figure 7G). Next, we compared the 10 most abundant species at the genus level. Notably, CAPE treatment markedly elevated the abundance of Bacteroides and decreased the abundance of Muribaculaceae and Parabacteroides (Figure 7H). Leveraging linear discriminant analysis of effect sizes (LEfSe), the microbiota was examined at the genus level to identify bacteria exhibiting substantial differences in abundance between the two cohorts (LDA score > 3.5). After CAPE treatment, the relative abundance of Bacteroides, Ruminococcus_gnavus_group, Muribaculum and Faecalibaculum OTUs notably increased, while that of Hungatella, Parasutterella and Muribaculaceae OTUs significantly decreased (Figure 7I). Bacteroides, Muribaculum and Faecalibaculum improve colonization resistance to C. difficile 26, 27, 28. Importantly, the abundance of these bacterial species increased following CAPE treatment (Figure 7J-L). In contrast, the abundance of Parasutterella, a bacterium that facilitates C. difficile colonization 29, was reduced in mice receiving CAPE (Figure 7M). Taken together, these findings suggest that CAPE treatment accelerates microbiota recovery during CDI.

CAPE alters the levels of gut metabolites in C. difficile-infected mice

In addition to the crucial role of the gut microbiota in conferring colonization resistance against C. difficile, gut microbiota-derived metabolites significantly contribute to the disease process 30. Metabolomic analysis was performed to evaluate the effects of CAPE treatment on gut metabolites in mice. The PLS- DA model exhibited significant separation of clusters between the model and CAPE treatment groups (Figure 8A). Among the 1599 metabolites, 109 showed substantial differences in levels between the two groups. Specifically, CAPE-treated mice exhibited 19 metabolites with upregulated expression and 50 metabolites with downregulated expression compared to those in the model group (fold change > 1.5) (Figure 8B). Furthermore, KEGG enrichment analysis revealed that the differentially abundant metabolites were significantly enriched in the purine metabolism and bile secretion pathways (Figure 8C). Adenosine is an endogenous purine nucleoside that binds to specific cell surface receptors to exert anti-inflammatory effects at high concentrations 31. Importantly, CAPE treatment of C. difficile-infected mice can increase the adenosine concentration in the gut (Figure 8D), which might be attributed to its inhibition of xanthine oxidase 32. Similarly, the levels of uric acid, the final product of purine metabolism, were markedly decreased in mice after receiving CAPE (Figure 8D). An earlier study reported that increasing the gut adenosine concentration with an adenosine deaminase inhibitor prevents C. difficile toxin A-induced enteritis in mice 33. Therefore, the increase in gut adenosine induced by CAPE might play an important role in its therapeutic efficacy in CDI.

Alterations in gut metabolites in CAPE-treated mice with CDI. (A)

The score plots show orthogonal partial least square discriminant analysis (PLS-DA) results. (B) Volcano plot representing the upregulated or downregulated gut metabolites in the CAPE treatment group compared to those in the model group. (C) Enrichment bubble map of the top 20 metabolic pathways in the CAPE treatment and model groups. (D) The 26 potential biomarkers are displayed in a heatmap, along with hierarchical clustering analysis.

C. difficile predominantly utilizes sugars and amino acids for energy production through various metabolic pathways, such as fermentation 34. Notably, Stickland fermentation, a Clostridia-specific type of amino acid fermentation, is associated with C. difficile pathogenesis 34. D-proline is one of the preferred nutrients used by C. difficile in the Stickland reaction and is critical for its growth 35, 36. Interestingly, treatment of C. difficile-infected mice with CAPE significantly reduced gut D-proline levels (Figure 8D). In addition, the metabolomics data also revealed a dramatic increase in the level of melatonin (Figure 8D), which is a ubiquitous indolamine that was previously shown to decrease the rate of recurrent CDI 37. Consistently, quantification of the melatonin content in mouse fecal supernatants using a melatonin detection kit revealed elevated levels of melatonin in CAPE-treated mice (Figure S3). Moreover, correlation analysis revealed a positive association between Faecalibaculum abundance and melatonin levels (Figure S4). Taken together, these results indicate that alterations in gut metabolites caused by CAPE treatment might contribute to its therapeutic efficacy in CDI.

Oral administration of melatonin alleviates the symptoms of CDI

Next, we investigated whether any of the upregulated metabolites had a direct influence on TcdB. To this end, we preincubated recombinant TcdB with each of the selected metabolites prior to addition to the cell cultures. Interestingly, treatment with TcdB and 100 μg/mL melatonin significantly alleviated the cell death induced by this toxin (Figure 9A). This observation prompted us to investigate the direct therapeutic potential of melatonin in CDI. Therefore, C. difficile-infected mice were orally administered 200 mg/kg melatonin at 24 h intervals. The weight changes and the dry/wet ratio of the stool samples were monitored daily for three days. The results showed that melatonin-treated mice displayed remarkable alleviation of diarrhea, as manifested by notable improvements in weight recovery and a significant increase in the dry/wet weight ratio on the 2nd and 3rd days after C. difficile challenge (Figure 9B-C). Consistent with these observations, histopathological assessment of colonic tissue sections also revealed attenuated colon damage upon melatonin treatment (Figure 9D-E). Taken together, these data indicate that a high dose of exogenous melatonin could protect mice from CDI and that the CAPE-induced increase in endogenous melatonin might improve the therapeutic outcome of patients with CAPE.

Oral administration of melatonin alleviates the pathology of CDI. (A)

The impact of upregulated metabolites on the cytotoxicity of TcdB. TcdB (0.2 ng/mL) was incubated with each of the metabolites (100 μg/mL) for 1 h at 37 °C before being added to the Vero cell culture. At 3 h after TcdB treatment, cell death was evaluated with a CCK-8 kit. (B) Body weight changes in melatonin- treated CDI mice (n=5). (C) The fecal dry/wet weight ratio in CDI mice after melatonin treatment (n=5). (D) Histopathological analysis of colon tissue from CDI mice treated with melatonin. (E) Histopathological scores of melatonin- treated CDI mice. * p <0.05; ** p <0.01 using one-way ANOVA.

Discussion

C. difficile has become the prevalent agent of nosocomial diarrhea and substantially threatens public health. In particular, the increase in morbidity and mortality caused by the prevalence of hypervirulent and drug-resistant strains has become a global concern 7. In 2019, C. difficile was listed by the US Centers for Disease Control and Prevention as an “urgent threat” that surpassed methicillin-resistant Staphylococcus aureus (MRSA) and ESBL-producing Enterobacterales 38. C. difficile is a commensal bacterium that colonizes the lower gastrointestinal tract 39. CDI typically occurs in susceptible individuals whose normal gut microbiota is disrupted by the administration of broad-spectrum antibiotics 39. Although CDI is considered an antibiotic-associated disease, the current treatment for CDI still relies on antimicrobial drugs, including metronidazole, fidaxomicin and vancomycin 9, 10. Despite antibiotic therapy, more than 35% of CDI patients are susceptible to recurrent infection, which is primarily attributable to the intrinsic effects of antimicrobial agents, which disrupt the microbiota 12. In addition, long-term administration of antibiotics promotes the selection of antibiotic-resistant subpopulations in the gut, thus increasing the risk of infections caused by antibiotic-resistant and opportunistic bacteria 40. As a result, the development of nonantibiotic therapeutics is needed for the treatment of CDI.

Indeed, numerous novel treatments for CDI are being investigated, among which the anti-virulence strategy represents one of the most promising and successful therapeutic approaches 22, 41. Unlike antibiotics, which target bacterial components that are critical for bacterial growth, anti-virulence drugs aim to disarm pathogenic bacteria by neutralizing their virulence factors 42, 43. Therefore, this strategy imposes less selective pressure that directs bacterial evolution of antimicrobial resistance 43. The disease progression of CDI is strictly dependent on large clostridial toxins (TcdA and TcdB), as strains lacking both toxins (i.e., A-B-) were completely avirulent and failed to induce any discernable disease symptoms in a hamster model of infection 44. Importantly, epidemic C. difficile strains with increased morbidity and mortality rates produce greater amounts of toxins 45. Based on these premises, TcdA and TcdB are considered ideal drug targets for CDI in the anti-virulence paradigm 46. Indeed, bezlotoxumab, a fully humanized monoclonal antibody that neutralizes TcdB, was recently approved by the FDA and used in CDI patients to reduce the rate of recurrence 47. This achievement has encouraged researchers to develop next-generation anti-virulence treatments. In particular, small molecule inhibitors of toxin production and/or activity have attracted increasing interest owing to their advantages in cost and the convenience of oral administration 41. The multidomain architecture and mechanisms of action of clostridial toxins have provided distinct options for blocking their toxicity. Indeed, previous efforts have identified diverse inhibitors that act on distinct steps of intoxication, such as the inhibition of toxin binding to cell receptors, the suppression of endosomal maturation, and the inactivation of GTD or CPD activities 23, 48, 49, 50, 51. Although which approach is the best remains unknown, we believe that an inhibitor that targets multiple steps of intoxication would be superior to those that act on a single step. In this study, we identified a series of caffeic acid derivatives that were potent inhibitors of TcdB via a phenotypic screen. Further mechanistic study revealed that CAPE, the phenethyl alcohol ester of caffeic acid, could block InsP6-induced autoproteolysis of TcdB and inactivate its GTD function (Figure 10). Moreover, CAPE treatment significantly protected against CDI in a mouse model of infection. Our results suggested that caffeic acid derivatives, particularly CAPE, are promising lead compounds for the development of anti- virulence drugs that target TcdB. CAPE is a major component of propolis, a natural bee-produced substance that has been used as a remedy for thousands of years in Europe and Asia 52. As a result, further investigations on the therapeutic potential of propolis against CDI, either in combination with standard therapy or as a single therapy, would be of great interest.

A model for CAPE-mediated suppression of TcdB cytotoxicity and protection of mice from CDI.

The intoxication of cells by TcdB involves multistep mechanisms. Direct interaction of CAPE with full-length TcdB blocks InsP6-induced autoproteolysis, resulting in decreased production of active GTD. The binding between GTD and CAPE inhibits its glucosyltransferase activity, thus decreasing the glucosylation level of Rac1. CAPE treatment of CDI mice markedly restored the diversity and composition of the gut microbiota and induced changes in gut metabolites, which might have contributed to the therapeutic outcomes of CDI following CAPE treatment.

Antibiotic-induced gut microbiota dysbiosis is a major risk factor for CDI 53. Thus, reconstituting the gut microbiota is a promising strategy for CDI therapy. Fecal microbiota transplantation (FMT) is considered an effective CDI therapeutic strategy and is widely used in clinical practice 54. Despite the high cure rate and increasing popularity of FMT, its safety remains controversial. Another prominent advantage of anti-virulence drugs is that they allow maintenance of the normal microbiota, which is of critical importance for the treatment of CDI. Importantly, CAPE treatment of C. difficile-infected mice restored the structure and composition of the microbiota (Figure 10). In particular, the abundance of Bacteroides, a critical gut microbiota component that is required for C. difficile resistance 26, was significantly elevated upon CAPE treatment. Interestingly, Bacteroides were also predominant in CDI patients administered RBX2660 (Rebyota), an FDA-approved fecal microbiota product 55, 56. The increased diversity of the gut microbiota might contribute to the therapeutic effects of CAPE in CDI.

In addition to the microbiota, gut metabolites also play crucial roles in the pathogenesis of CDI, either by promoting or suppressing disease progression 30. For example, bile acids and amino acids promote spore germination and pathogenesis, while secondary bile acid and short-chain fatty acids aid in colonization resistance and suppress spore germination 30. In addition to restoration of the gut microbiota, metabolomics analysis of mouse fecal samples revealed dramatic alterations in gut metabolites upon CAPE treatment (Figure 10). The differentially abundant metabolites were significantly enriched in the purine metabolism and bile metabolism pathways. Remarkably, the level of adenosine, an endogenous purine nucleoside that is released from cells upon injury or inflammation, was dramatically increased in CAPE-treated mice. In contrast, the amount of uric acid, the final product of purine metabolism, was reduced. Previous studies showed that the increase in adenosine concentrations caused by the inhibition of adenosine deaminase via EHNA prevents C. difficile TcdA-mediated damage and inflammation 33. This protection might be a result of the anti-inflammatory effects of adenosine, which rely on the activation of its receptors 57. In addition, we observed a reduction in the amount of D-proline, which is one of the preferred amino acids utilized by C. difficile to generate energy through Stickland reactions 35. Melatonin is a ubiquitous indolamine that has broad biological activities 58. Notably, we detected a significantly increase in the amount of melatonin in the feces of mice that were administered CAPE. Importantly, an earlier study demonstrated a decreased risk of recurrent CDI in patients exposed to melatonin, possibly owing to its antibacterial and anti-inflammatory activities 37. In this study, we found that oral administration of exogenous melatonin alleviated the symptoms of CDI in a mouse model of infection. Although determining the exact mechanism will require further investigation, we propose that the melatonin- mediated inhibition of TcdB intoxication observed in this study might contribute to its therapeutic effect. Taken together, these results indicate that the alterations in gut metabolites caused by CAPE might also contribute to treatment outcomes for CDI patients.

Methods

Cell lines

All cell lines (HeLa, Caco-2, Vero, HEK293T and THP-1) were obtained from ATCC (Manassas, VA). HeLa, Caco-2, Vero and HEK293T cells were cultured in Dulbecco’s modified Eagle’s medium (DMEM) (Invitrogen, Carlsbad, CA) supplemented with 10% fetal bovine serum (FBS) (Invitrogen) and 1% (w/v) penicillin‒streptomycin. THP-1 cells were cultured in Roswell Park Memorial Institute (RPMI) 1640 medium (Invitrogen) supplemented with 10% FBS and 1% (w/v) penicillin‒streptomycin. All cell lines were cultured at 37 °C in a 5% CO2 atmosphere.

Bacterial strains and reagents

C. difficile BAA-1870 (kindly provided by Aiwu Wu, Guangzhou Medical University) was cultured in brain-heart infusion broth (BHI). The compounds were identified from two main sources: FDA-approved drugs obtained from ApexBio Technology (Houston, TX, USA) and small molecule natural compounds obtained from Chengdu Ruifensi Biotechnology Co., Ltd. (Chengdu, China). LysoTracker Red DND-99, kanamycin, and cefoperazone sodium were purchased from Dalian Meilun Biotechnology Co., Ltd. (Dalian, China).

Mice

Male C57BL/6 mice aged 6-8 weeks were purchased from Liaoning Changsheng Biotechnology Co., Ltd., and housed in a specific pathogen-free (SPF) facility. The mice lived in a controlled environment with a temperature of 23±2 °C and a humidity of 55±10 °C. Prior to infection, the mice were acclimated to the environment for a minimum of one week. All mouse experiments were approved by the Jilin University Institutional Animal Care Committee, Jilin University and strictly conducted according to the guidelines of this committee (SY202309036).

Protein expression and purification

The expression and purification of full-length TcdB were conducted following the method described by Yang et al. 59. In brief, Bacillus megaterium harboring pHis1522-TcdB was inoculated into LB media supplemented with a final concentration of 10 μg/mL tetracycline. Once the OD600nm reached approximately 0.3, xylose (0.5% w/v) was added to induce protein expression at 37 °C for 16 h. Then, 100 mL of the bacterial cells was harvested by centrifugation, resuspended in 5 mL of lysis buffer containing 20 mM imidazole, 300 mM NaCl, 20 mM NaH2PO4, 500 μM EDTA, and protease inhibitor cocktail (adjusted to pH 8.0), and subjected to sonication. The cell lysate was then centrifuged at 13,400 × g for 20 min to remove cell debris. The TcdB protein was purified using nickel column affinity chromatography. The purity of the protein was confirmed by SDS‒PAGE, and the protein was subsequently dialyzed against buffer containing 10% glycerol, 20 mM Tris (pH 7.5), and 150 mM NaCl. After the protein concentration was determined, the purified TcdB protein was aliquoted and stored at -80 °C for further use.

For the purification of GTD, amplified DNA products were inserted into pET28a (Novagen), and the Escherichia coli (E. coli) strain BL21 (DE3) was used for the expression and purification of GTD. Transformed E. coli pET28a- GTD was cultured in Luria–Bertani (LB) medium supplemented with 10 μg/mL kanamycin. Once the OD600 reached 0.6, the expression of proteins was induced with 0.2 mM isopropyl-β-d-thiogalactoside (IPTG) overnight at 18 °C. The bacterial cells were then harvested by centrifugation and subjected to sonication. The cell lysate was then centrifuged at 13,400 × g for 20 min to remove cell debris. GTD protein was purified using nickel affinity chromatography and dialyzed in buffer containing 10% glycerol, 20 mM Tris (pH 7.5), and 150 mM NaCl. The GTD mutants were introduced into pET28a-GTD using the Quick Change Site-Directed Mutagenesis Kit (Stratagene, San Diego, CA, USA) with the primers listed in Table S2. The subsequent purification was performed using the same procedure as described for wild-type GTD.

Cell rounding assay

Vero or HeLa cells were seeded in 96-well sterile plates using serum-free DMEM overnight, and the compounds were preincubated with TcdB for 1 h. After incubation, the TcdB-compound complex was added to cells for 1 h at 37 °C. The cell morphology was visualized under a light microscope.

LysoTracker assay

The pH of endosomes was determined by LysoTracker Red DND-99. Caco-2 or THP-1 cells in complete DMEM were seeded in 96-well sterile plates overnight. The media was changed to serum-free DMEM for 1 h, and CAPE was preincubated with TcdB for 1 h, followed by addition to the cells and incubation for 2 h. Baf-A1 served as a control. After incubation, LysoTracker and Hoechst dyes were added and the cells were incubated for 1 h. Excess dye was removed by changing the media, and alterations in lysosomal red intensity were documented using a fluorescence microscope.

In vitro TcdB autoprocessing assay

The self-cleavage ability of TcdB primarily depends on its cysteine protease activity. CAPE (1-16 μg/mL) was preincubated with TcdB for 1 h, followed by the addition of InsP6 and incubation at 37 °C for 6 h. Subsequently, loading buffer was added for SDS‒PAGE analysis, and the results were visualized using Coomassie Brilliant Blue staining.

Rac1 glucosylation assay

HEK293T cells were seeded into a 6-well plate. The next day, TcdB and different concentrations of CAPE (4-32 μg/mL) were added to cells. After 2 h, the cells were harvested, lysed using lysis buffer, and subjected to centrifugation at 13,400 × g for 10 min to collect the supernatant. The collected supernatant was then mixed with loading buffer and subjected to SDS‒PAGE followed by western blot analysis. The glycosylation level of Rac1 was evaluated by determining the levels of non-glycosylated Rac1 (BD Biosciences, Mississauga, ON) and total Rac1 (Abcam).

UDP-GloTM UDP-glucose hydrolase assay

GTD activity was assessed according to the manufacturer’s instructions. A total volume of 16 μL, consisting of 100 nM GTD, various concentrations of CAPE and glucosylation buffer, was prepared. Four microliters of UDP-glucose at a final concentration of 25 μM was added to the mixture, which was then allowed to react at room temperature for 15 min. To stop the reaction, 10 μL of the reaction mixture was removed and placed in a white polystyrene 96-well plate containing 10 μL of UDP detection reagent. The reaction took place at room temperature for 1 h. The luminescence value was measured at a 750 ms integral, and the GTD activity was determined based on these measurements.

Circular dichroism (CD) spectroscopy

CD spectra of TcdB, TcdB-CAPE, GTD, and GTD-CAPE were obtained using a CD spectrophotometer (MOS-500; Bio-Logic, France). TcdB or GTD at a concentration of 0.2 mg/mL was incubated with 32 μg/mL CAPE. Changes in the secondary structure of TcdB or GTD were assessed at room temperature using a quartz cuvette with an optical path length of 1 mm. The scan wavelength range was set from 190 to 250 nm with a resolution of 0.2 nm and a bandwidth of 1 nm. The secondary structures of TcdB, TcdB-CAPE, GTD, and GTD-CAPE were analyzed using the BeStSel web server.

Molecular docking

CAPE was obtained as a ligand from PubChem, and its topology was generated using AmeberTools. The crystal structure of GTD was retrieved from the Protein Data Bank (PDB) and processed using AutoDockTools. Docking calculations were performed using AutodockVina, and molecular dynamics simulations were carried out using Gromacs 2020.6. The RMSD of the receptor and ligand was monitored to assess conformational changes during the binding process. The interaction energy was analyzed using the Poisson-Boltzmann surface area (MM-PBSA) method, which is a molecular mechanics approach.

Flow cytometry

Flow cytometry was used to quantify the cellular uptake of TcdB via endocytosis. To facilitate detection, full-length TcdB was labeled with Alexa Fluor®488 NHS Ester (Yeasen Biotechnology Co., Ltd., Shanghai, China). Prior to cell treatment, Alexa Fluor 488-TcdB was preincubated with CAPE at a concentration of 32 μg/mL for 30 min. The treated solution was then added to Caco-2 cells and incubated for 1 h at 37 °C. Following three washes with phosphate-buffered saline (PBS), the Caco-2 cells were collected, and the fluorescence signal was measured using flow cytometry.

Cytotoxicity assay

Vero or THP-1 cells were plated into 96-well plates. The next day, different concentrations of CAPE were prepared by diluting them in DMEM and adding them to the cells. After incubation at 37 °C for 3 h, the viability of cells was assessed using a Cell Counting Kit-8 (CCK-8) (Beyotime Biotechnology Co., Shanghai, China) according to the manufacturer’s instructions.

Biacore analysis

The affinity analysis of TcdB or GTD with CAPE was conducted using a Biacore T100 device (Biacore Inc., Uppsala, Sweden) according to the manufacturer’s protocol. A CM5 chip (General Electric Company, GE) was utilized to immobilize the TcdB or GTD proteins. Various concentrations of CAPE were injected, and all steps were performed in running buffer consisting of 5% DMSO in PBS-P20 (GE). Evaluation version 3.1 software provided by Biacore and the 1:1 Langmuir combined model were employed for data analysis.

Mouse infection model

For the acute toxicity model, after pretreatment of TcdB with CAPE (2 or 4 μg/mL) for 1 h, the mice were intraperitoneally injected with a complex of TcdB- solvent or TcdB-CAPE (30 ng of TcdB per mouse). The survival of mice was observed for 72 h.

For the CDI model, mice were given cefoperazone (0.5 g/L) in their drinking water for 5 days (from day -7 to day -2), followed by switching to sterilized water for 2 days. On day 0, the mice were challenged with 1 × 108 colony-forming units (CFUs) of C. difficile BAA-1870 (Ribotype 027) by oral gavage. For the CAPE or melatonin treatment group, mice were treated with CAPE (30 mg/kg by oral gavage) or melatonin (200 mg/kg by oral gavage) every day after C. difficile challenge. The weight loss and fecal state of the mice were monitored daily. Fecal samples were collected on days 0-3 to evaluate the stool water content as a marker of diarrhea. Additionally, fecal samples on day 3 were collected to assess C. difficile colonization. Colonic tissues were collected on day 3 post infection, stained with hematoxylin and eosin (H&E) and scored as previously described 60.

Fecal samples were collected on day 3 post infection, frozen in liquid nitrogen and transferred to the laboratory with dry ice for 16S rRNA sequencing and untargeted metabolomics analysis (Novogene Bioinformatics Technology Co., Ltd., Tianjin, China). The melatonin level in mouse stool supernatant was determined using the Mouse MT (Melatonin) ELISA Kit (Sangon Biotechnology Co., Ltd., Shanghai, China).

Statistical analysis

Western blot densitometry and LysoTracker fluorescence intensity were measured using ImageJ software (NIH, Bethesda, MD, USA). Statistical analysis was performed using GraphPad Prism 9.0 (GraphPad Software, USA). Unpaired two-tailed Student’s t tests were used for comparing two groups, and one-way analysis of variance (ANOVA) was used for analyzing data from more than three groups. The statistical significance was determined using the log- rank (Mantel–Cox) test (survival rates) and Mann–Whitney U test (tissue bacterial load) in vivo. All the data are expressed as the mean ± standard deviation (SD). * indicates p <0.05 and ** indicates p <0.01.

eLife Assessment

The manuscript by Guo and colleagues reports valuable findings about the inhibitory activity of caffeic acid phenethyl ester (CAPE) against TcdB, a key toxin produced by Clostridioides difficile. C. difficile infections are a major public health concern, and this manuscript provides interesting data on toxin inhibition by CAPE, a potentially promising therapeutic alternative for this disease. The strength of the evidence to support the conclusions is solid, with some concerns about the moderate effects on the mouse infection model and direct binding assays of CAPE to the toxin.

Reviewer #1 (Public review):

Summary:

In this manuscript, Guo and colleagues used a cell rounding assay to screen a library of compounds for inhibition of TcdB, an important toxin produced by Clostridioides difficile. Caffeic acid and derivatives were identified as promising leads, and caffeic acid phenethyl ester (CAPE) was further investigated.

Strengths:

Considering the high morbidity rate associated with C. difficile infections (CDI), this manuscript presents valuable research in the investigation of novel therapeutics to combat this pressing issue. Given the rising antibiotic resistance in CDI, the significance of this work is particularly noteworthy. The authors employed a robust set of methods and confirmatory tests, which strengthened the validity of the findings. The explanations provided are clear, and the scientific rationale behind the results is well-articulated. The manuscript is extremely well-written and organized. There is a clear flow in the description of the experiments performed. Also, the authors have investigated the effects of CAPE on TcdB in careful detail and reported compelling evidence that this is a meaningful and potentially useful metabolite for further studies.

Weaknesses:

This is really a manuscript about CAPE, not caffeic acid, and the title should reflect that. Also, a few details are missing from the description of the experiments. The authors should carefully revise the manuscript to ascertain that all details that could affect the interpretation of their results are presented clearly. Just as an example, the authors state in the results section that TcdB was incubated with compounds and then added to cells. Was there a wash step in between? Could compound carryover affect how the cells reacted independently from TcdB? This is just an example of how the authors should be careful with descriptions of their experimental procedures. Lastly, authors should be careful when drawing conclusions from the analysis of microbiota composition data. Ascribing causality to correlational relationships is a recurring issue in the microbiome field. Therefore, I suggest authors carefully revise the manuscript and tone down some statements about the impact of CAPE treatment on the gut microbiota.

Reviewer #2 (Public review):

Summary:

This work is towards the development of nonantibiotic treatment for C. difficile. The authors screened a chemical library for activity against the C. difficile toxin TcdB, and found a group of compounds with antitoxin activity. Caffeic acid derivatives were highly represented within this group of antitoxin compounds, and the remaining portion of this work involves defining the mechanism of action of caffeic acid phenethyl ester (CAPE) and testing CAPE in mouse C. difficile infection model. The authors conclude CAPE attenuates C. difficile disease by limiting toxin activity and increasing microbial diversity during C. difficile infection.

Strengths/ Weaknesses:

The strategy employed by the authors is sound although not necessarily novel. A compound that can target multiple steps in the pathogenies of C. difficile would be an exciting finding. However, the data presented does not convincingly demonstrate that CAPE attenuates C. difficile disease and the mechanism of action of CAPE is not convincingly defined. The following points highlight the rationale for my evaluation.

(1) The toxin exposure in tissue culture seems brief (Figure 1). Do longer incubation times between the toxin and cells still show CAPE prevents toxin activity?

(2) The conclusion that CAPE has antitoxin activity during infection would be strengthened if the mouse was pretreated with CAPE before toxin injections (Figure 1D).

(3) CAPE does not bind to TcdB with high affinity as shown by SPR (Figure 4). A higher affinity may be necessary to inhibit TcdB during infection. The GTD binds with millimolar affinity and does not show saturable binding. Is the GTD the binding site for CAPE? Autoprocessing is also affected by CAPE indicating CAPE is binding non-GTD sites on TcdB.

(4) In the infection model, CAPE does not statistically significantly attenuate weight loss during C. difficile infection (Figure 6). I recognize that weight loss is an indirect measure of C. difficile disease but histopathology also does not show substantial disease alleviation (see below).

(5) In the infection model (Figure 6), the histopathology analysis shows substantial improvement in edema but limited improvement in cellular infiltration and epithelial damage. Histopathology is probably the most critical parameter in this model and a compound with disease-modifying effects should provide substantial improvements.

(6) The reduction in C. difficile colonization is interesting. It is unclear if this is due to antitoxin activity and/or due to CAPE modifying the gut microbiota and metabolites (Figure 6). To interpret these data, a control is needed that has CAPE treatment without C. difficile infection or infection with an atoxicogenic strain.

(7) Similar to the CAPE data, the melatonin data does not display potent antitoxin activity and the mouse model experiment shows marginal improvement in the histopathological analysis (Figure 9). Using 100 µg/ml of melatonin (~ 400 micromolar) to inactivate TcdB in cell culture seems high. Can that level be achieved in the gut?

(8) The following parameters should be considered and would aid in the interpretation of this work. Does CAPE directly affect the growth of C. difficile? Does CAPE affect the secretion of TcdB from C. difficile? Does CAPE alter the sporulation and germination of C. diffcile?

Reviewer #3 (Public review):

Summary:

The study is well written, and the results are solid and well demonstrated. It shows a field that can be explored for the treatment of CDI

Strengths:

The results are really good, and the CAPE shows a good and promising alternative for treating CDI. The methodology and results are well presented, with tables and figures that corroborate them. It is solid work and very promising.

Weaknesses:

Some references are too old or missing.

Author response:

The following is the authors’ response to the original reviews

Public Reviews:

Reviewer #1 (Public review):

Summary:

In this manuscript, Guo and colleagues used a cell rounding assay to screen a library of compounds for inhibition of TcdB, an important toxin produced by Clostridioides difficile. Caffeic acid and derivatives were identified as promising leads, and caffeic acid phenethyl ester (CAPE) was further investigated.

Strengths:

Considering the high morbidity rate associated with C. difficile infections (CDI), this manuscript presents valuable research in the investigation of novel therapeutics to combat this pressing issue. Given the rising antibiotic resistance in CDI, the significance of this work is particularly noteworthy. The authors employed a robust set of methods and confirmatory tests, which strengthened the validity of the findings. The explanations provided are clear, and the scientific rationale behind the results is well-articulated. The manuscript is extremely well-written and organized. There is a clear flow in the description of the experiments performed. Also, the authors have investigated the effects of CAPE on TcdB in careful detail and reported compelling evidence that this is a meaningful and potentially useful metabolite for further studies.

Weaknesses:

This is really a manuscript about CAPE, not caffeic acid, and the title should reflect that. Also, a few details are missing from the description of the experiments. The authors should carefully revise the manuscript to ascertain that all details that could affect the interpretation of their results are presented clearly. Just as an example, the authors state in the results section that TcdB was incubated with compounds and then added to cells. Was there a wash step in between? Could compound carryover affect how the cells reacted independently from TcdB? This is just an example of how the authors should be careful with descriptions of their experimental procedures. Lastly, authors should be careful when drawing conclusions from the analysis of microbiota composition data. Ascribing causality to correlational relationships is a recurring issue in the microbiome field. Therefore, I suggest authors carefully revise the manuscript and tone down some statements about the impact of CAPE treatment on the gut microbiota.

Thanks for your constructive suggestion. We have carefully revised the manuscript, including the description of title, results and methods sections.

Reviewer #2 (Public review):

Summary:

This work is towards the development of nonantibiotic treatment for C. difficile. The authors screened a chemical library for activity against the C. difficile toxin TcdB, and found a group of compounds with antitoxin activity. Caffeic acid derivatives were highly represented within this group of antitoxin compounds, and the remaining portion of this work involves defining the mechanism of action of caffeic acid phenethyl ester (CAPE) and testing CAPE in mouse C. difficile infection model. The authors conclude CAPE attenuates C. difficile disease by limiting toxin activity and increasing microbial diversity during C. difficile infection.

Strengths/ Weaknesses:

The strategy employed by the authors is sound although not necessarily novel. A compound that can target multiple steps in the pathogenies of C. difficile would be an exciting finding. However, the data presented does not convincingly demonstrate that CAPE attenuates C. difficile disease and the mechanism of action of CAPE is not convincingly defined. The following points highlight the rationale for my evaluation.

(1) The toxin exposure in tissue culture seems brief (Figure 1). Do longer incubation times between the toxin and cells still show CAPE prevents toxin activity?

Thanks for your comments. The cytotoxicity assay was employed to directly assess the protective capacity of CAPE against cell death induced by TcdB. Our observations at 1 and 12 h post-TcdB exposure revealed that CAPE effectively mitigated the toxic effects of the TcdB at both time points, demonstrating its potent protective role. Please see Figure S1.

(2) The conclusion that CAPE has antitoxin activity during infection would be strengthened if the mouse was pretreated with CAPE before toxin injections (Figure 1D).

Thanks for your constructive comments. According to your suggestion, we administered TcdB 2 h after pretreatment with CAPE. The outcomes demonstrated that CAPE pretreatment significantly enhanced the survival rate of the intoxicated mice, confirming that CAPE retains its antitoxin efficacy during the infection process. Please see Figure S2.

(3) CAPE does not bind to TcdB with high affinity as shown by SPR (Figure 4). A higher affinity may be necessary to inhibit TcdB during infection. The GTD binds with millimolar affinity and does not show saturable binding. Is the GTD the binding site for CAPE? Auto processing is also affected by CAPE indicating CAPE is binding non-GTD sites on TcdB.

Thanks for your comments. Our findings indicate that the GTD domain is a critical binding site for CAPE. CAPE exerts its protective effects at multiple stages of TcdB-mediated cell death, including inhibiting TcdB's self-cleavage and blocking the activity of GTD, thereby preventing the glycosylation modification of Rac1 by TcdB.

(4) In the infection model, CAPE does not statistically significantly attenuate weight loss during C. difficile infection (Figure 6). I recognize that weight loss is an indirect measure of C. difficile disease but histopathology also does not show substantial disease alleviation (see below).

Thanks for your comments. Our comparative analysis revealed a notable distinction in the body weight of mice on the third day post-infection (Figure 6B). Similarly, the dry/wet stool ratio exhibited a comparable pattern, suggesting that treatment with phenethyl caffeic acid ameliorated Clostridium difficile-induced diarrhea to a significant degree (Figure 6C).

(5) In the infection model (Figure 6), the histopathology analysis shows substantial improvement in edema but limited improvement in cellular infiltration and epithelial damage. Histopathology is probably the most critical parameter in this model and a compound with disease-modifying effects should provide substantial improvements.

Thanks for your comments. Edema, inflammatory factor infiltration, and epithelial damage served as key evaluation metrics. Statistical analysis revealed that the pathological scores of mice treated with CAPE were markedly reduced compared to those in the model group (Figure 6F).

(6) The reduction in C. difficile colonization is interesting. It is unclear if this is due to antitoxin activity and/or due to CAPE modifying the gut microbiota and metabolites (Figure 6). To interpret these data, a control is needed that has CAPE treatment without C. difficile infection or infection with an atoxicogenic strain.

The observed reduction in C. difficile fecal colonization following drug treatment may be attributed to the CAPE's antitoxin properties or its capacity to modify the intestinal microbiota and metabolites. These two mechanisms likely work in tandem to combat CDI. CDI is primarily triggered by the toxins A (TcdA) and B (TcdB) secreted by the bacterium. Certain therapies, including monoclonal antibodies like bezlotoxumab, target CDI by neutralizing these toxins, thereby mitigating gut damage and subsequent C. difficile colonization(1,2). The establishment of C. difficile in the gut is intricately linked to the equilibrium of the intestinal microbiota. Although antibiotic treatments can inhibit C. difficile growth, they may also disrupt the microbial balance, potentially facilitating the overgrowth of other pathogens. Consequently, interventions such as fecal microbiota transplantation (FMT) are designed to reestablish gut flora balance and consequently decrease C. difficile colonization(3,4). Moreover, the administration of probiotics and prebiotics is considered to reduce C. difficile colonization by modifying the gut environment(5,6).

(7) Similar to the CAPE data, the melatonin data does not display potent antitoxin activity and the mouse model experiment shows marginal improvement in the histopathological analysis (Figure 9). Using 100 µg/ml of melatonin (~ 400 micromolar) to inactivate TcdB in cell culture seems high. Can that level be achieved in the gut?

The uptake and dissemination of melatonin within the body varies with the dose administered. For instance, in rats, the bioavailability of melatonin following administration was found to be 53.5%, whereas in dogs, bioavailability was nearly complete (100%) at a dose of 10 mg/kg, yet it decreased to 16.9% at a lower dose of 1 mg/kg(7). This data suggests that the absorption of melatonin differs across various animal species and is influenced by the dose administered. Moreover, it underscores the higher potential bioavailability of melatonin, implying that a dose of 200 mg/kg should be adequate to achieve the desired concentration in the body post-administration.

(8) The following parameters should be considered and would aid in the interpretation of this work. Does CAPE directly affect the growth of C. difficile? Does CAPE affect the secretion of TcdB from C. difficile? Does CAPE alter the sporulation and germination of C. diffcile?

We incorporated CAPE into the MIC assay for detecting C. difficile, as well as for assessing the sporulation capacity of C. difficile and evaluating the secretion level of TcdB. The findings revealed that CAPE markedly repressed tcdB transcription at a concentration of 16 μg/mL and effectively suppressed the growth and sporulation of C. difficile BAA-1870 at a concentration of 32 μg/mL. Please see Figure S3.

References:

(1) Skinner AM, et al. Efficacy of bezlotoxumab to prevent recurrent Clostridioides difficile infection (CDI) in patients with multiple prior recurrent CDI. Anaerobe. 2023 Dec; 84: 102788.

(2) Wilcox MH, et al. Bezlotoxumab for Prevention of Recurrent Clostridium difficile Infection. N Engl J Med. 2017 Jan 26;376(4):305-317.

(3) Khoruts A, Sadowsky MJ. Understanding the mechanisms of faecal microbiota transplantation. Nat Rev Gastroenterol Hepatol. 2016 Sep;13(9):508-16.

(4) Khoruts A, Staley C, Sadowsky MJ. Faecal microbiota transplantation for Clostridioides difficile: mechanisms and pharmacology. Nat Rev Gastroenterol Hepatol. 2021 Jan;18(1):67-80.

(5) Mills JP, Rao K, Young VB. Probiotics for prevention of Clostridium difficile infection. Curr Opin Gastroenterol. 2018 Jan;34(1):3-10.

(6) Lau CS, Chamberlain RS. Probiotics are effective at preventing Clostridium difficile-associated diarrhea: a systematic review and meta-analysis. Int J Gen Med. 2016 Feb 22; 9:27-37.

(7) Yeleswaram K, et al. Pharmacokinetics and oral bioavailability of exogenous melatonin in preclinical animal models and clinical implications. J Pineal Res. 1997 Jan;22(1):45-51.

Reviewer #3 (Public review):

Summary:

The study is well written, and the results are solid and well demonstrated. It shows a field that can be explored for the treatment of CDI.

Strengths:

The results are really good, and the CAPE shows a good and promising alternative for treating CDI. The methodology and results are well presented, with tables and figures that corroborate them. It is solid work and very promising.

Weaknesses:

Some references are too old or missing.

Thanks for your constructive suggestion. We have included and refreshed several references to enhance the manuscript.

Recommendations for the authors:

Reviewer #1 (Recommendations for the authors):

While the manuscript convincingly demonstrates that CAPE affects the TcdB toxin and reduces its toxicity in vitro, it would be beneficial to include data on the effect of CAPE on the growth of C. difficile. This would help ensure that the observed in vivo effects are not merely due to reduced bacterial growth but rather due to the specific action of CAPE on the toxin.

Thanks for your constructive suggestion. We have augmented our findings with the impact of CAPE on the bacteria themselves, revealing that CAPE not only hampers the growth of the bacterial cells but also suppresses their capacity to produce spores. Please see Figure S3.

(1) Line 41, line 115 - authors should clarify what they mean when mentioning Bacteroides within parentheses.

Thanks for your comments. We have completed the corresponding modifications according to the suggestions.

(2) Line 71 - Is C. difficile really found "in the environment"?

Thanks for your comments. C. difficile is prevalent across various natural settings, including soil and water ecosystems. A study has identified highly diverse strains of this bacterium within environmental samples(1). Moreover, the significant presence of C. difficile in soil and lawn specimens collected near Australian hospitals indicates that the organism is indeed a common inhabitant in the environment(2).

(3) Lines 128-130 - Was there a wash step here? What could be the impact of compound carryover in this experiment?

Thanks for your comments. Following pre-incubation of TcdB with CAPE, remove the compounds that have not bound to TcdB through centrifugation. The persistence of the compound in the culture post-washing could result in an inflated assessment of its efficacy, particularly if it continues to engage with TcdB or the cells beyond the initial 1-hour pre-incubation window. The carryover of the compound might also give rise to misleading positive results, where the compound seems to confer protection or inhibition against TcdB-mediated cell rounding, whereas such effects are actually due to the lingering activity of the compound. This carryover could skew the determination of the compound's minimum effective concentration, as the effective concentration interacting with the cells might be inadvertently elevated. Furthermore, if the compounds possess cytotoxic properties or impact cell viability, carryover could generate artifacts in cell morphology that are unrelated to the direct interaction between TcdB and the compounds.

(4) Lines 133-134 - I suggest authors mention how many caffeic acid derivatives there were in the entire library so that the suggested "enrichment" of them in the group of bioactive compounds can be better judged.

Thanks for your comments. The natural compound library contained eight caffeic acid derivatives, of which methyl caffeic acid and ferulic acid displayed no efficacy. This information has been incorporated into the manuscript.

(5) Line 135 - I recommend the authors add the molarity of the compound solutions used.

Thanks for your comments. We have completed the corresponding modifications according to the suggestions.

(6) Line 247 - I think the term "CAPE mice" is confusing. Please use a full description.

Thanks for your comments. We have completed the corresponding modifications according to the suggestions.

(7) Line 248 - I also think the terms "model mice" and "model group" are confusing. Maybe call them "control mice"?

Thanks for your comments. The terms "model mice" and "model group" are indeed synonymous, and we have subsequently clarified that control mice refer to those that have not been infected with C. difficile.

(8) Line 273 - "most abundant species at the genus level" is incorrect. I think what you mean is "most abundant TAXA".

Thanks for your comments. We have completed the corresponding modifications according to the suggestions.

(9) Line 278 - Please include your p-value cut-off together with the LDA score.

Thanks for your comments. We have revised the above description to “LDA score > 3.5, p < 0.05”.

(10) Line 292 - Details on how metabolomics was performed should be included here.

Thanks for your comments. We have completed the corresponding modifications according to the suggestions.

(11) Line 299 - 1.5 is a fairly low cut-off. The authors should at a minimum also include the p-value cut-off used.

Response: Thanks for your comments. We have revised the above description to “fold change > 1.5, p < 0.05”.

(12) Line 307 - Purine "degradation" would be better here.

Thanks for your comments. We have completed the corresponding modifications according to the suggestions.

(13) Line 328 onward - The melatonin experiment is a weird one. Although I fully understand the rationale behind testing the effect of melatonin in the mouse model, the idea that just because melatonin levels changed in the gut it would act as a direct inhibitor of TcdB was very far-fetched, even though it ended up working. Authors should explain this in the manuscript.

Thanks for your comments. Furthermore, beyond our murine studies, we have confirmed that melatonin significantly diminishes TcdB-induced cytotoxicity at the cellular level (Figure 9A). Additionally, it has been documented that melatonin, acting as an antimicrobial adjuvant and anti-inflammatory agent, can decrease the recurrence of CDI(3). Consequently, we contend that the aforementioned statement is substantiated.

(14) Lines 429-435 - There are seemingly contradictory pieces of information here. The authors state that adenosine is released from cells upon inflammation and that CAPE treatment caused an increase in adenosine levels. Later in this section, the authors state that adenosine prevents TcdA-mediated damage and inflammation. This should be clarified and better discussed.

Thanks for your comments. Adenosine modulates immune responses and inflammatory cascades by interacting with its receptors, including its capacity to suppress the secretion of specific pro-inflammatory mediators. We have updated this depiction in the manuscript.

(15) Lines 513-514 - How was this phenotype quantified?

Thanks for your comments. Initially, we introduced TcdB at a final concentration of 0.2 ng/mL along with various concentrations of compounds into 1 mL of medium for a 1-h pre-incubation period. Subsequently, unbound compounds were removed through centrifugation, and the resulting mixture was then applied to the cells.

(16) Figure 3 - panels are labeled incorrectly.

Thanks for your comments. We have completed the corresponding modifications according to the suggestions.

(17) Figure 5C - it is unclear what the different colors and labels represent.

Thanks for your comments. In the depicted graph, blue denotes the total binding energy, red signifies the electrostatic interactions, green corresponds to the van der Waals forces, and orange indicates solvation or hydration effects. The horizontal axis represents the mutation of the amino acid residue at the respective position to alanine. As illustrated in Figure 5C, the mutations W520A and GTD exhibit the highest binding energies.

References:

(1) Janezic S, et al. Highly Divergent Clostridium difficile Strains Isolated from the Environment. PLoS One. 2016 Nov 23;11(11): e0167101.

(2) Perumalsamy S, Putsathit P, Riley TV. High prevalence of Clostridium difficile in soil, mulch and lawn samples from the grounds of Western Australian hospitals. Anaerobe. 2019 Dec; 60:102065.

(3) Sutton SS, et al. Melatonin as an Antimicrobial Adjuvant and Anti-Inflammatory for the Management of Recurrent Clostridioides difficile Infection. Antibiotics (Basel). 2022 Oct 25;11(11):1472.

Reviewer #2 (Recommendations for the authors):

Minor comments and questions.

(1) Which form of TcdB is being used in these experiments?

Thanks for your comments. The TcdB proteins used in this study are TcdB1 subtypes.

(2) Why are THP-1 cells being used in these assays?

Thanks for your comments. For the purposes of this study, we employed a diverse array of cell lines, including Vero, HeLa, THP-1, Caco-2, and HEK293T. Each cell line was selected to serve a specific experimental objective. The inclusion of the THP-1 cell line was necessitated by the need to incorporate a macrophage cell line to ensure the comprehensive nature of our experiments, allowing for the testing of both epithelial cells and macrophages. C. difficile is a kind of intestinal pathogenic bacteria, and immune clearance plays a vital role in the process of pathogen infection, so THP-1 cells are used as important immune cells.

(3) Please improve the quality of the microscopy images in Figure 1.

Thanks for your comments. We have improved the quality of the microscopy images in Figure 1.

(4) Does the flow cytometry experiment in Figure 2B show internalization? Surface-bound toxins would provide the same histogram.

Thanks for your comments. Figure 2B was employed to assess the internalization of TcdB, and the findings indicate that CAPE does not influence the internalization process of TcdB.

(5) The sensogram in Figure 4A does not look typical and should be clarified.

Thanks for your comments. Typically, small molecules and proteins engage in a rapid binding and dissociation dynamic. However, as depicted in Figure 4A, the interaction between CAPE and TcdB demonstrates a gradual progression towards equilibrium. This behavior can be primarily explained by the swift occupation of the protein's primary binding sites by the small molecule in the initial stages. Subsequently, CAPE binds to secondary or lower affinity sites, extending the time needed to reach equilibrium. Additionally, the likelihood of CAPE binding to multiple sites on TcdB requires time for the exploration and occupation of these diverse locations before equilibrium is attained, we have incorporated an analysis of this potential scenario into the manuscript.

Reviewer #3 (Recommendations for the authors):

These are my suggestions for the text:

(1) Line 29: high recurrent rates.

Thanks for your comments. We have completed the corresponding modifications according to the suggestions.

(2) Line 32: Where is the caffeic acid identified? I think a line should be included.

Thanks for your comments. Caffeic acid was identified from natural compounds library and we have completed the corresponding modifications according to the suggestions.

(3) Line 39: C. difficile is not italic.

Thanks for your comments. We have completed the corresponding modifications according to the suggestions.

(4) Line 41: Bacteroides spp.

Thanks for your comments. We have completed the corresponding modifications according to the suggestions.

(5) Line 56: This number of casualties 56.000 is still happening or it was in the past?

Thanks for your comments. The mortality rates reported in the manuscript reflect a downturn in the incidence and fatality of CDI around 2017(1), as the infection gained broader recognition. Nonetheless, a recent study reveals that the mortality rate for CDI cases in Germany can soar to 45.7% within a year, with the overall economic burden amounting to approximately 1.6 billion euros. This underscores the ongoing significance of CDI as a global public health challenge(2).

(6) Line 104: Where did the idea of testing caffeic acid come from? Any previous study of the authors? Any studies with the inhibition of other pathogens?

Thanks for your comments. Initially, we conducted a screen of a compound library comprising 2,076 compounds and identified several potent inhibitors, which, upon structural analysis, were revealed to be caffeic acid derivatives. Prior to our investigation, no studies had explored the potential of CAPE in this context.

(7) Line 115: Bacteroides spp.

Thanks for your comments. We have completed the corresponding modifications according to the suggestions.

Results section

(8) Did the authors try the caffeic acid with the TcdA or binary toxin? I know this is not the purpose of the study, but TcdA toxin has a high identity structure with TcdB and generates inflammation in the gut via neutrophils. Negative strains for the major toxins and positive for the binary toxin also cause severe cases of CDI.

Thanks for your comments. Although we acknowledge the significance of TcdA and binary toxins in CDI, we did not investigate the impact of CAPE on these toxins. Our focus was exclusively on the effect of CAPE against TcdB, as it is the primary virulence factor in C. difficile pathogenesis. Since TcdA and TcdB are highly similar in structure, we will analyze the neutralization effect of CAPE on TcdA in later studies.

(9) Does caffeic acid have any effect on C. difficle? Or does it only gain the toxins? That would be ideal.

Thanks for your comments. We have included additional related assays in our study. Beyond directly neutralizing TcdB, CAPE also demonstrates the capacity to inhibit the growth and spore formation of C. difficile.

(10) Line 230: C. difficile BAA-1870 is a clinical strain? There are no details about it in the paper.

Thanks for your comments. C. difficile BAA-1870 (RT027/ST1), a highly virulent isolate frequently employed in research(3-6), was kindly donated by Professor Aiwu Wu. We have meticulously noted the PCR ribotype in our manuscript.

(11) Line 236: Did the mice fully recover from CDI after the administration of the CAPE? Was one dose enough?

Thanks for your comments. CAPE was administered orally at 24 h intervals, commencing with the initial dose on Day 0. By the time a significant difference was observed on Day 3, the treatment had been administered a total of three times.

Methodology

(12) Most of the methods do not have a reference.

Thanks for your comments. We have added several references to the methods.

Discussion section

(13) The first two paragraphs of the discussion should be summarized. Those details were already explained in the introduction.

Thanks for your comments. The discussion section and the introduction address slightly different focal points; therefore, we aim to retain the first two paragraphs to maintain continuity and context.

(14) Line 382: Bezolotoxumab was approved by the FDA in 2016. It is not recent.

Thanks for your comments. We have revised the above description.

(15) Line 410: "Despite the high 410 cure rate and increasing popularity of FMT, its safety remains controversial. Although this is true, recently (2022) the FDA approved the Rebyota, which was later cited by the authors.

Thanks for your comments. We have revised the above description.

(16) Lines 415-416: "the abundance of Bacteroides, a critical gut microbiota component that is required for C. difficile resistance". There is only one reference cited by the authors. I suppose that if it is true, more studies should be mentioned. Why are probiotics with Bacteroides spp. not available in the market?

Thanks for your comments. We have supplemented additional references. The scarcity of probiotic products containing Bacteroides spp. on the market is primarily attributable to the stringent requirements of their survival conditions. As most Bacteroides spp. are anaerobic, they thrive in oxygen-deprived environments. This unique survival trait poses challenges in maintaining their viability during product preservation and distribution, which in turn escalates production costs and complexity. Furthermore, despite the significant role of Bacteroides in gut health, research into its potential probiotic benefits and safety is comparatively underexplored.

References:

(1) Guh AY, et al. Emerging Infections Program Clostridioides difficile Infection Working Group. Trends in U.S. Burden of Clostridioides difficile Infection and Outcomes. N Engl J Med. 2020 Apr 2;382(14):1320-1330.

(2) Schley K, et al. Costs and Outcomes of Clostridioides difficile Infections in Germany: A Retrospective Health Claims Data Analysis. Infect Dis Ther. 2024 Nov 20.

(3) Saito R, et al. Hypervirulent clade 2, ribotype 019/sequence type 67 Clostridioides difficile strain from Japan. Gut Pathog. 2019 Nov 4; 11:54.

(4) Pellissery AJ, Vinayamohan PG, Venkitanarayanan K. In vitro antivirulence activity of baicalin against Clostridioides difficile. J Med Microbiol. 2020 Apr;69(4):631-639.

(5) Shao X, et al. Chemical Space Exploration around Thieno[3,2-d]pyrimidin-4(3H)-one Scaffold Led to a Novel Class of Highly Active Clostridium difficile Inhibitors. J Med Chem. 2019 Nov 14;62(21):9772-9791.

(6) Mooyottu S, Flock G, Venkitanarayanan K. Carvacrol reduces Clostridium difficile sporulation and spore outgrowth in vitro. J Med Microbiol. 2017 Aug;66(8):1229-1234.

Version 2

Abstract

Abstract

Clostridioides difficile infection (CDI) is the leading cause of hospital-acquired diarrhea that seriously threatens public health. The disruption of normal gut microbiota by the use of broad-spectrum antimicrobial agents enables C. difficile to proliferate in the colon. The emergence and prevalence of hypervirulent C. difficile strains result in increased morbidity, mortality, and high recurrence rates of CDI, thus creating a pressing need for novel therapeutics. The multi-domain toxins TcdA and TcdB are the primary determinants of CDI pathogenesis, rendering them ideal drug targets in the anti-virulence paradigm. In this study, we identified caffeic acid and its derivatives from natural compounds library as active inhibitors of TcdB via a cell-based high-throughput phenotypic screening. Further mechanistic investigations revealed that caffeic acid phenethyl ester (CAPE) could directly bind to TcdB, thus suppressing InsP6-induced autoproteolysis and inhibiting glucosyltransferase activity. CAPE treatment remarkably reduces the pathology of CDI in a murine infection model in terms of alleviated diarrhea symptoms, decreased bacterial colonization, and relieved histopathological lesions. Moreover, CAPE treatment of C. difficile-challenged mice induces a remarkable increase in the diversity and composition of the gut microbiota (e.g., Bacteroides spp.) and alterations of gut metabolites (e.g., adenosine, D-proline, and melatonin), which might partially contribute to the therapeutic outcomes of CAPE against CDI. Our results reveal the potential of CAPE as a therapeutic for the management of CDI, or CAPE might serve as a lead compound for the development of antivirulence drugs targeting TcdB.

Introduction

Clostridioides difficile (C. difficile) is an obligate anaerobic, spore-forming, gram-positive bacterial pathogen that is the most common cause of hospital-acquired and antibiotic-associated gastrointestinal diseases, including mild diarrhea and life-threatening pseudomembranous colitis 1, 2, 3. In the United States, C. difficile is responsible for almost half a million infections and approximately 29,000 deaths per year, resulting in an estimated annual healthcare-associated cost of up to $4.8 billion 4, 5. Importantly, the incidence, severity, mortality and healthcare costs of C. difficile infection (CDI) have dramatically increased in the 21st century, at least partially owing to the frequent use of antimicrobial agents and the emergence of hypervirulent strains (e.g., PCR ribotypes 027 and 078) 6, 7, 8. Limited options are available for the treatment of CDI, while antibiotics including metronidazole, fidaxomicin and vancomycin, are the mainstays of clinical practice 9, 10. However, antibiotic treatment is ineffective in more than 35% of CDI patients and can lead infection recurrence 11, 12. Moreover, the rapid evolution of antibiotic resistance in C. difficile has further compromised the therapeutic efficacy of current treatment options 13. Therefore, the high morbidity and mortality, high recurrence rates, and emergence of multiple antibiotic-resistant isolates have made this pathogen a major threat to public health; these factors highlight the urgent need for novel therapies.

C. difficile widely exists in the environment and in the intestinal tract of humans and animals and is mainly transmitted via the fecal-oral route 14. The pathogenesis of CDI is correlated with the use of antibiotics, which impair the structure and composition of the normal gut microbiota, thus allowing C. difficile to proliferate in the colon 14. Therefore, previous antibiotic therapy is the major risk factor for the development of CDI. In the colon, the dormant spores of C. difficile germinate into oxygen-sensitive vegetative cells that produce toxins and cause symptoms of disease 15. C. difficile secretes two large clostridial exotoxins (TcdA and TcdB), which are directly responsible for the clinical pathology of CDI 16, 17. Epidemiological studies suggest that most disease-causing C. difficile strains have genes that encode TcdB 18, 19. TcdA and TcdB have identical functional domains: a N-terminal glucosyltransferase catalytic domain (GTD), a cysteine protease active domain (CPD), a transmembrane domain (DRBD), and a C-terminal receptor binding domain (CROPs) 20. When secreted into the colon, both toxins are internalized by host cells through receptor-mediated, clathrin-dependent endocytosis. Following a decrease in endosomal pH, the toxins undergo conformational changes in the DRBD, resulting in insertion of the toxin into the endosome membrane, pore formation, and the release of the GTD and CPD domains into the cytosol. Then, CPD binding with cytosolic inositol hexakisphosphate (InsP6) activates its protease activity, leading to the cleavage and release of the GTD. Eventually, the GTD utilizes UDP-glucose to glucosylate Rho and/or Ras family GTPases, thus disrupting GTPase signaling and resulting in deleterious effects, including cytopathic “rounding”, dysregulation of the actin cytoskeleton, apoptosis, and pyroptosis1. Although both toxins are cytotoxic to host cells, TcdB shows greater toxicity than TcdA 21.

Disruption of the gut microbiota by antimicrobial agents results in high recurrence rates, which limits the application of antimicrobial therapy for CDI 14. Consequently, there is a critical need to develop nonantibiotic treatments for CDI, either as standalone or augmented therapies. Importantly, strategies targeting bacterial virulence factors, including toxins, are termed anti-virulence strategies and have attracted interest in the 21st century 22. Owing to its essential role in C. difficile pathogenesis, TcdB has become an ideal target for the development of anti-virulence drugs 23. In this study, caffeic acid and a series of its derivatives were found to inhibit TcdB-mediated cell rounding via cell-based screening. Among these potential inhibitors, caffeic acid phenethyl ester (CAPE) had the greatest inhibitory effect. Mechanistically, CAPE bound to both full-length TcdB and the GTD fragment, leading to alterations in the secondary structure. This interaction suppressed the InsP6-mediated self-cleavage of TcdB, as well as the enzymatic activity of GTD, thereby decreasing the level of glucosylated Rac1 in TcdB-treated host cells and further inhibiting the cytopathology and cytotoxicity of TcdB. Oral administration of CAPE significantly protected against CDI in models of animal infection. In addition to direct toxin inhibition, the therapeutic effect may also be attributed to CAPE-induced alteration of the gut microbiota (e.g., Bacteroides spp.) as well as gut metabolites (e.g., adenosine, D-proline, and melatonin). Overall, our study revealed that caffeic acid and its derivatives, particularly CAPE, may serve as promising lead compounds for the development of anti-virulence drugs to treat CDI.

Results

CAPE inhibits cell rounding induced by TcdB

TcdB can inactivate Rho GTPases via glucosylation, leading to the loss of cytoskeletal structure and cell-rounding. Consequently, we utilized this phenotype to identify potential TcdB inhibitors from 2076 chemicals, including 1515 FDA-approved drugs and 561 natural compounds. The minimal concentration of recombinant His6-TcdB (0.2 ng/mL) that induced 100% cell rounding within 1 h was used for drug screening. This concentration of TcdB was preincubated with 8 μg/mL of the individual compounds for 1 h, the compounds that were not bound to TcdB were removed and then added to added to Vero cell cultures. At 1 h post-TcdB treatment, cell morphology was visualized under a light microscope (Figure 1A). These analyses allowed us to identify 22 candidate compounds with potent inhibitory effects (Figure 1B). Interestingly, six out of the 22 compounds were caffeic acid derivatives. The natural compound library contained eight caffeic acid derivatives, of which methyl caffeic acid and ferulic acid displayed no efficacy. Caffeic acid itself only slightly inhibited TcdB-induced cell rounding, with a 50% inhibitory concentration (IC50) of 527.5 μg/mL (1.86 mM) (Table S1). Notably, except for rosmarinic acid, the IC50 values for the other screened caffeic acid derivatives were less than 6 μg/mL, suggesting that these compounds potently inhibited TcdB-mediated toxicity (Table S1). CAPE is the phenethyl alcohol ester of caffeic acid and a main active component of propolis with broad pharmacological properties, such as anticancer, anti-inflammatory, antiviral, antibacterial, antifungal, antioxidant, and immunomodulatory activities 24. As CAPE has the lowest IC50 (3.0 μg/mL) among the active caffeic acid derivatives, it was selected as the representative compound for subsequent investigations (Table S1). After preincubation with increasing concentrations of CAPE, dose-dependent inhibition of cell rounding was observed in both Vero and HeLa cells (Figure 1C). The protective influence of CAPE against TcdB-induced Vero cell death remained evident even when the treatment duration was prolonged to 12 h (Figure S1A-B).

CAPE suppresses TcdB-mediated cell rounding.

(A) Schematic of the process used for the identification of small molecules that inhibit the cell rounding phenotype induced by TcdB. TcdB (0.2 ng/mL) was preincubated with 8 μg/mL of individual compounds prior to addition to the cell cultures. (B) Active compounds were identified from the chemical library consisting of 2076 chemicals. The red dots represent caffeic acid and its derivatives. (C) Dose-dependent CAPE inhibition of TcdB-induced rounding of Vero (upper) and HeLa (lower) cells. Scale bar, 100 μm. The dose titration curves are presented on the right. (D) The survival rates of mice that were intraperitoneally injected with CAPE-pretreated or native TcdB (30 ng). * p <0.05 and ** p <0.01 by log-rank (Mantel-Cox) test.

In addition, we also determined whether inhibition of TcdB by CAPE could lead to loss of toxin function in vivo using a murine systemic intoxication model. A lethal dose of recombinant His6-TcdB (30 ng) was preincubated with either DMSO or CAPE (2 or 4 μg/mL) for 1 h at 37 °C, followed by intraperitoneal injection into mice. Notably, all mice in the DMSO group (n=10) died within 36 h after toxin injection. However, 100% of mice pretreated with CAPE 4 (μg/mL) toxin survived at the monitoring endpoint (72 h) (n=10) (Figure 1D). Furthermore, we evaluated the efficacy of CAPE following an intraperitoneal injection by administering it 2 h prior to the injection of TcdB, in order to gauge the survival rate of the mice. The outcomes revealed that a 2 h pretreatment with CAPE at doses of 30 and 50 mg/kg significantly enhanced the mice’s protection rate by 57.1% (Figure S2). Simultaneously, we assessed the impact of CAPE on C.difficile itself. CAPE markedly repressed tcdB gene expression at a concentration of 16 μg/mL (Figure S3A) and inhibited the growth and spore formation of C. difficile BAA-1870 at a concentration of 32 μg/mL (Figure S3B-C). Taken together, these data indicate that CAPE is a potent toxin inhibitor of TcdB and may serve as a promising antimicrobial agent.

CAPE decreases InsP6-induced autoproteolysis of TcdB and inhibits the enzymatic activity of GTD

TcdB is a multi-domain containing toxin that induces toxicity in host cells via an intricate and multistep mechanism. This mechanism involves (i) receptor-mediated endocytosis, (ii) pore formation and subsequent membrane translocation, (iii) autoproteolysis and release of GTD, and (iv) the glucosylation of small GTPases (Figure 2A). To determine the mechanism by which CAPE suppresses the cytotoxicity of TcdB, we first labeled full-length TcdB with Alexa Fluor 488; then, the concentration of the endocytosed toxin was determined via flow cytometry. Apparently, incubation of CAPE with TcdB did not significantly affect its entry into host cells (Figure 2B). Next, we stained CAPE-treated cells with LysoTracker, a fluorescent dye that accumulates in the acidic organelles of live cells. In contrast to the vacuolar-type H+-ATPase (V-ATPase) inhibitor bafilomycin A1 (Baf-A1), CAPE treatment did not alter the intensity of LysoTracker fluorescence staining in either THP-1 or Caco-2 cells, suggesting that inactivation of TcdB by CAPE is not due to changes in the pH of endosomal compartments (Figure 2C-D and Figure S4A-B).

Mechanism by which CAPE suppresses TcdB intoxication.

(A) Schematic illustration of the multistep mechanism of TcdB intoxication of host cells. (B) Determination of internalized TcdB levels. Alexa Fluor 488-labeled TcdB was incubated with 32 μg/mL CAPE for 30 min before it was added to the cell culture. The cellular uptake of TcdB was measured by flow cytometry analysis. (C) Lysosomal activity in host cells treated with CAPE was measured via LysoTracker staining. The nuclei were stained with DAPI. Bar, 100 μm. (D) The intensities of LysoTracker staining were quantified using ImageJ. * p <0.05; ** p <0.01 using one-way ANOVA. (E-F) CAPE inhibits InsP6-induced autoprocessing of full-length TcdB. Two micrograms of TcdB was preincubated with increasing concentrations of CAPE for 1 h. Self-cleavage of TcdB was initiated by the addition of 3 μg of InsP6 and the cells were incubated for 6 h at 37 °C. Reactions were terminated with 5x SDS loading buffer, and the release of GTD was visualized by CBB staining of the SDS-PAGE gel (E). The level of released GTD in panel E was quantified by ImageJ (F) * p <0.05; ** p <0.01 using one-way ANOVA. (G) Dose-dependent inhibition of the GTD activity of TcdB by CAPE. Recombinant GTD of TcdB was mixed with different concentrations of CAPE (0-32 μg/mL) and incubated for 15 min at 25 °C. GTD activity was measured by the UDP-Glo™ glycosyltransferase assay. (H-I) Intracellular Rac1 glucosylation in cells treated with TcdB. Native or CAPE-pretreated TcdB was added to the cell culture and cells were incubated for 2 h. The cells were lysed, and the glucosylation of Rac1 was analyzed by western blotting using antibodies recognizing total Rac1 and unglycosylated Rac1 (H). The level of Rac1 glucosylation shown in H was quantified by ImageJ (I). * p <0.05; ** p <0.01 using one-way ANOVA.

These observations indicate that CAPE might exert its effects after membrane translocation. Indeed, preincubation of full-length TcdB with CAPE blocked the InsP6-induced autoproteolysis of TcdB, as revealed by the gradual decrease in the production of released GTP, followed by the addition of increasing amounts of CAPE in vitro (Figure 2E-F). Furthermore, we determined the glucosyltransferase activity using recombinant GTD in the absence or presence of different doses of CAPE. Strikingly, incubation with 2 μg/mL CAPE resulted in a significant reduction in GTD activity, whereas 32 μg/mL CAPE compromised 87.5% of the enzymatic activity (Figure 2G). Importantly, in response to suppression of InsP6-induced autoproteolysis and GTD function, treatment of TcdB-challenged cells with CAPE markedly decreased the level of glucosylated Rac1 in a dose-dependent manner (Figure 2H-I). Taken together, these results suggest that CAPE disrupts InsP6-induced autoproteolysis of TcdB and inhibits the enzymatic activity of GTD, thereby decreasing the glucosylation of Rac1 and alleviating cell rounding.

CAPE alters the secondary structure of full-length TcdB and GTD

To further clarify the mechanisms used by CAPE to inactivate TcdB, we determined the secondary structures of full-length TcdB and GTD in the absence or presence of CAPE by circular dichroism spectroscopy. Apparently, incubation of CAPE with full-length TcdB resulted in a significant increase in the α-helix content (15.4% to 21.0%) and a concomitant decrease in the β-pleated sheet content (34.5% to 20.9%) (Figure 3A-B). Similarly, the secondary structure of GTD was markedly altered upon incubation with CAPE, with the proportion of α-helices decreasing from 37.9% to 0.0% and the proportion of β-pleated sheets decreasing from 37.2% to 10.9% (Figure 3C-D). Taken together, these data suggest that CAPE may interact with full-length TcdB and GTD to induce conformational changes, thus impairing their functions.

CAPE alters the secondary structure of TcdB and GTD.

(A-B) The secondary structure of TcdB (A) or GTD (B) in the presence or absence of 32 μg/mL CAPE was determined by CD spectroscopy. (C-D) Calculated CD spectra of CAPE-treated TcdB (C) and GTD (D). The wavelength for CD spectroscopy was set at 190-250 nm.

CAPE directly interacts with full-length TcdB and GTD

To investigate the potential interaction between CAPE and TcdB or GTD, the binding constants were determined by surface plasmon resonance (SPR) analysis. The results showed that CAPE has strong binding affinity with full-length TcdB and GTD, and the KD values were 1.269×10-5 M and 1.007×10-3 M, respectively (Figure 4A-B). The interaction between CAPE and TcdB exhibits a slow approach to equilibrium, which may primarily be attributed to the rapid occupation of the protein’s primary binding sites by the small molecule during the initial phase. Following this, CAPE proceeds to bind to sites with secondary or reduced affinity, a process that prolongs the time required to achieve equilibrium. Furthermore, the possibility that CAPE binds to multiple sites on TcdB necessitates a duration for the exploration and occupation of these various sites before equilibrium is established. These data demonstrate the direct interaction between CAPE and full-length TcdB/GTD.

Direct interaction of CAPE with TcdB and GTD.

(A-B) Full-length TcdB or GTD was immobilized on a CM5 chip, followed by the injection of various concentrations of CAPE. The response unit (RU) value and affinities of CAPE interactions with TcdB (A) and GTD (B) were determined by a Biacore T100.

Next, we performed molecular docking and molecular simulation dynamics analyses to investigate the interaction between CAPE and GTD (Figure 5A). The results revealed that Leu265, Asp286, Thr465, Ile466, Leu519, and Trp520 in GTD were potential binding sites for CAPE (Figure 5A). The root mean square deviation (RMSD) of the GTD/CAPE complex was stable over a 100 ns period of simulation (Figure 5B). Further energy decomposition analysis showed that the side chains of Leu265, Thr465, Leu519, and Trp520 of GTD contributed to binding with CAPE via van der Waals interactions, among which Trp520 had the highest binding energy (Figure 5C). Moreover, we mutated each of the residues that were involved in binding and subsequently determined their glucosyltransferase activities in the presence or absence of CAPE. The enzymatic activities of these mutants were similar to that of wild-type GTD (Figure 5D). Mutations in GTDL265A, GTDT465A, GTDI466A, and GTDL519A did not affect the inhibitory effect of CAPE (Figure 5E-I). However, the inhibition of GTDW520A and GTDD286A by CAPE was notably compromised (Figure 5J-K).

Asp286 and Trp520 in the GTD are critical for CAPE-mediated inhibition of GTD.

(A) Three-dimensional structure determination of GTD with the CAPE complex was performed by molecular modeling. (B) The RMSD values of the GTD-CAPE complex. (C) Decomposition of the binding energy on a per-residue basis in the binding sites of the GTD-CAPE complex. (D) Glucosyltransferase activity of wild-type GTD and the GTD mutants. (E-K) Residual activity of wild-type GTD (E) and its mutants GTDL265A (F), GTDT465A (G), GTDI466A (H), GTDL519A (I), GTDW520A (J) and GTDD286A (K) in the presence of increasing concentrations of CAPE (0-4 μg/mL). * p <0.05; ** p <0.01 using one-way ANOVA.

CAPE treatment reduces the pathology of CDI in a murine model

The abovementioned observation that CAPE suppresses the function of TcdB prompted us to further evaluate the potential therapeutic effects of this compound in CDI. To this end, we employed a well-established murine model in which infected mice can develop symptoms resembling those observed in human CDIs (25). First, cytotoxicity measurements revealed that the viability of both THP-1 and Vero cells was not affected by CAPE treatment (Figure S5A-B), thus indicating low toxicity. Mice were preconditioned with drinking water containing 0.5 g/L cefoperazone for 5 days, followed by normal drinking water for 2 days. After challenge with C. difficile BAA-1870 via gavage, mice were then treated daily with either vehicle or 30 mg/kg CAPE by gavage throughout the disease course (day 0 to day 5) (Figure 6A). Severe weight loss in C. difficile-infected mice was observed on days 2 and 3 post infection and was accompanied by diarrhea (Figure 6B). Notably, compared to the infection control group, CAPE treatment significantly increased the dry/wet weight ratio of the mouse feces, suggesting the alleviation of diarrhea symptoms (Figure 6C). Correspondingly, the weight loss of mice was relieved upon receiving CAPE (Figure 6B). Moreover, fecal pellets collected on day 3 were used to determine the bacterial burden. The results revealed that CAPE treatment significantly decreased the number of C. difficile CFUs (Figure 6D). On day 3 post-C. difficile infection, the colon was collected for histopathological analysis. Remarkably, C. difficile-infected mice exhibited conspicuous colon edema, epithelial cell injury and inflammatory factor infiltration (Figure 6E). In contrast, these histopathological lesions were significantly less common in mice administered CAPE (Figure 6F). Taken together, these data demonstrate the efficacy of CAPE in the treatment of CDI.

CAPE treatment blocks the pathology of CDI in mice.

(A) Schematic of the process used for the establishment of the CDI mouse model. (B) Weight changes in CDI mice after CAPE or vehicle treatment (n=10). (C) The fecal dry/wet weight ratio in CDI mice treated with CAPE or vehicle (n=6). * p <0.05; ** p <0.01 using one-way ANOVA. (D) Bacterial colonization in the feces of CDI mice treated with CAPE or vehicle (n=15). * p <0.05; ** p <0.01 using the Mann_Whitney U test. (E) Pathological analysis of colon sections from CDI mice treated with CAPE or vehicle. The black boxes are the regions that were magnified and are shown underneath. (F) Histopathological scores of vehicle-treated and CAPE-treated mice in terms of cell infiltration (orange), epithelial damage (purple), and edema (green). * p <0.05; ** p <0.01 using one-way ANOVA.

CAPE induces distinct microbiota changes in CAPE-treated mice when compared with model mice

Antibiotic-induced dysbiosis of the gut microbiota promotes the development of CDI 25. Therefore, we next determined whether CAPE has any impact on gut microbial communities that may contribute to its therapeutic efficacy. To this end, on the 3rd day after C. difficile challenge, fecal samples from the model group and CAPE treatment group were collected and subjected to microbiota analysis via 16S rRNA gene sequencing. Remarkably, as reflected by the Venn diagram, CAPE treatment resulted in significant changes in the microbial composition compared to those in the model group, and only 16.42% of the operational taxonomic units (OTUs) were shared (Figure 7A). In addition, both the α– and β-diversity of the gut microbiota in CAPE-treated mice were significantly greater than those in model mice. Specifically, the increase in the Chao1 richness estimator of mice in the CAPE treatment group suggested a greater number of microbial species and increased evenness in the fecal samples (Figure 7B-C); the increase in the Shannon-Wiener diversity index and the Simpson diversity index reflected a significant increase in microbiota diversity (Figure 7D-E). Principal coordinate analysis (PCoA) revealed distinct confidence ellipses between the bacterial communities of the CAPE-treated group and the model group, thereby revealing substantial differences in their microbiota profiles (Figure 7F).

The influence of CAPE treatment on the diversity and composition of the gut microbiota in CDI mice.

(A) Venn diagram illustrating the numbers of differentially abundant bacteria identified using a single model and their cooperation. (B-E) Quantification of the α diversity of the gut microbiota according to the observed species richness (B-C), Shannon index (D), and Simpson index (E). (F) PCoA showing β diversity based on the weighted UniFrac distance. (G-H) Relative abundance of the top 10 predominant bacteria classified at the phylum (G) and genus (H) levels. (I) Significant changes in flora among the model and CAPE groups, as measured by LEfSe analysis (LDA score [log 10] > 3.5). (J-M) Relative abundances of Bacteroides (J), Muribaculum (K), Faecalibaculum (L), and Parasutterella (M) in mice in the model and CAPE treatment groups. * p <0.05; ** p <0.01 using an unpaired Student’s t test.

A comparative analysis was performed to examine changes at the phylum level following CAPE treatment. Bacteroidota, Firmicutes, and Proteobacteria were the major constituents in both the CAPE treatment and model groups, suggesting that the administration of CAPE did not significantly alter the mouse microbiota at the phylum level (Figure 7G). Next, we compared the 10 most abundant TAXA at the genus level. Notably, CAPE treatment markedly elevated the abundance of Bacteroides and decreased the abundance of Muribaculaceae and Parabacteroides (Figure 7H). Leveraging linear discriminant analysis of effect sizes (LEfSe), the microbiota was examined at the genus level to identify bacteria exhibiting substantial differences in abundance between the two cohorts (LDA score > 3.5, p < 0.05). After CAPE treatment, the relative abundance of Bacteroides, Ruminococcus_gnavus_group, Muribaculum and Faecalibaculum OTUs notably increased, while that of Hungatella, Parasutterella and Muribaculaceae OTUs significantly decreased (Figure 7I). Bacteroides, Muribaculum and Faecalibaculum improve colonization resistance to C. difficile 26, 27, 28. Importantly, the abundance of these bacterial species increased following CAPE treatment (Figure 7J-L). In contrast, the abundance of Parasutterella, a bacterium that facilitates C. difficile colonization 29, was reduced in mice receiving CAPE (Figure 7M). Taken together, these findings suggest that CAPE treatment accelerates microbiota recovery during CDI.

CAPE alters the levels of gut metabolites in C. difficile-infected mice

In addition to the crucial role of the gut microbiota in conferring colonization resistance against C. difficile, gut microbiota-derived metabolites significantly contribute to the disease process 30. Metabolomic analysis was performed to evaluate the effects of CAPE treatment on gut metabolites in mice. On the 3rd day after infection, the feces of mice in each group were frozen by liquid nitrogen for metabolomics analysis. The PLS-DA model exhibited significant separation of clusters between the model and CAPE treatment groups (Figure 8A). Among the 1599 metabolites, 109 showed substantial differences in levels between the two groups. Specifically, CAPE-treated mice exhibited 19 metabolites with upregulated expression and 50 metabolites with downregulated expression compared to those in the model group (fold change > 1.5, p < 0.05) (Figure 8B). Furthermore, KEGG enrichment analysis revealed that the differentially abundant metabolites were significantly enriched in the purine metabolism and bile secretion pathways (Figure 8C). Adenosine is an endogenous purine nucleoside that binds to specific cell surface receptors to exert anti-inflammatory effects at high concentrations 31. Importantly, CAPE treatment of C. difficile-infected mice can increase the adenosine concentration in the gut (Figure 8D), which might be attributed to its inhibition of xanthine oxidase 32. Similarly, the levels of uric acid, the final product of purine metabolism, were markedly degraded in mice after receiving CAPE (Figure 8D). An earlier study reported that increasing the gut adenosine concentration with an adenosine deaminase inhibitor prevents C. difficile toxin A-induced enteritis in mice 33. Therefore, the increase in gut adenosine induced by CAPE might play an important role in its therapeutic efficacy in CDI.

Alterations in gut metabolites in CAPE-treated mice with CDI.

(A) The score plots show orthogonal partial least square discriminant analysis (PLS-DA) results. (B) Volcano plot representing the upregulated or downregulated gut metabolites in the CAPE treatment group compared to those in the model group. (C) Enrichment bubble map of the top 20 metabolic pathways in the CAPE treatment and model groups. (D) The 26 potential biomarkers are displayed in a heatmap, along with hierarchical clustering analysis.

C. difficile predominantly utilizes sugars and amino acids for energy production through various metabolic pathways, such as fermentation 34. Notably, Stickland fermentation, a Clostridia-specific type of amino acid fermentation, is associated with C. difficile pathogenesis 34. D-proline is one of the preferred nutrients used by C. difficile in the Stickland reaction and is critical for its growth 35, 36. Interestingly, treatment of C. difficile-infected mice with CAPE significantly reduced gut D-proline levels (Figure 8D). In addition, the metabolomics data also revealed a dramatic increase in the level of melatonin (Figure 8D), which is a ubiquitous indolamine that was previously shown to decrease the rate of recurrent CDI 37. Consistently, quantification of the melatonin content in mouse fecal supernatants using a melatonin detection kit revealed elevated levels of melatonin in CAPE-treated mice (Figure S6). Moreover, correlation analysis revealed a positive association between Faecalibaculum abundance and melatonin levels (Figure S7). Taken together, these results indicate that alterations in gut metabolites caused by CAPE treatment might contribute to its therapeutic efficacy in CDI.

Oral administration of melatonin alleviates the symptoms of CDI

Next, we investigated whether any of the upregulated metabolites had a direct influence on TcdB. To this end, we preincubated recombinant TcdB with each of the selected metabolites prior to addition to the cell cultures. Interestingly, treatment with TcdB and 100 μg/mL melatonin significantly alleviated the cell death induced by this toxin (Figure 9A). This observation prompted us to investigate the direct therapeutic potential of melatonin in CDI. Therefore, C. difficile-infected mice were orally administered 200 mg/kg melatonin at 24 h intervals. The weight changes and the dry/wet ratio of the stool samples were monitored daily for three days. The results showed that melatonin-treated mice displayed remarkable alleviation of diarrhea, as manifested by notable improvements in weight recovery and a significant increase in the dry/wet weight ratio on the 2nd and 3rd days after C. difficile challenge (Figure 9B-C). Consistent with these observations, histopathological assessment of colonic tissue sections also revealed attenuated colon damage upon melatonin treatment (Figure 9D-E). Taken together, these data indicate that a high dose of exogenous melatonin could protect mice from CDI and that the CAPE-induced increase in endogenous melatonin might improve the therapeutic outcome of patients with CAPE.

Oral administration of melatonin alleviates the pathology of CDI.

(A) The impact of upregulated metabolites on the cytotoxicity of TcdB. TcdB (0.2 ng/mL) was incubated with each of the metabolites (100 μg/mL) for 1 h at 37 °C before being added to the Vero cell culture. At 3 h after TcdB treatment, cell death was evaluated with a CCK-8 kit. (B) Body weight changes in melatonin-treated CDI mice (n=5). (C) The fecal dry/wet weight ratio in CDI mice after melatonin treatment (n=5). (D) Histopathological analysis of colon tissue from CDI mice treated with melatonin. (E) Histopathological scores of melatonin-treated CDI mice. * p <0.05; ** p <0.01 using one-way ANOVA.

Discussion

C. difficile has become the prevalent agent of nosocomial diarrhea and substantially threatens public health. In particular, the increase in morbidity and mortality caused by the prevalence of hypervirulent and drug-resistant strains has become a global concern 7. In 2019, C. difficile was listed by the US Centers for Disease Control and Prevention as an “urgent threat” that surpassed methicillin-resistant Staphylococcus aureus (MRSA) and ESBL-producing Enterobacterales 38. C. difficile is a commensal bacterium that colonizes the lower gastrointestinal tract 39. CDI typically occurs in susceptible individuals whose normal gut microbiota is disrupted by the administration of broad-spectrum antibiotics 39. Although CDI is considered an antibiotic-associated disease, the current treatment for CDI still relies on antimicrobial drugs, including metronidazole, fidaxomicin and vancomycin 9, 10. Despite antibiotic therapy, more than 35% of CDI patients are susceptible to recurrent infection, which is primarily attributable to the intrinsic effects of antimicrobial agents, which disrupt the microbiota 12. In addition, long-term administration of antibiotics promotes the selection of antibiotic-resistant subpopulations in the gut, thus increasing the risk of infections caused by antibiotic-resistant and opportunistic bacteria 40. As a result, the development of nonantibiotic therapeutics is needed for the treatment of CDI.

Indeed, numerous novel treatments for CDI are being investigated, among which the anti-virulence strategy represents one of the most promising and successful therapeutic approaches 22, 41. Unlike antibiotics, which target bacterial components that are critical for bacterial growth, anti-virulence drugs aim to disarm pathogenic bacteria by neutralizing their virulence factors 42, 43. Therefore, this strategy imposes less selective pressure that directs bacterial evolution of antimicrobial resistance 43. The disease progression of CDI is strictly dependent on large clostridial toxins (TcdA and TcdB), as strains lacking both toxins (i.e., A-B-) were completely avirulent and failed to induce any discernable disease symptoms in a hamster model of infection 44. Importantly, epidemic C. difficile strains with increased morbidity and mortality rates produce greater amounts of toxins 45. Based on these premises, TcdA and TcdB are considered ideal drug targets for CDI in the anti-virulence paradigm46. Indeed, bezlotoxumab, a fully humanized monoclonal antibody that neutralizes TcdB, was approved by the FDA and used in CDI patients to reduce the rate of recurrence 47. This achievement has encouraged researchers to develop next-generation anti-virulence treatments. In particular, small molecule inhibitors of toxin production and/or activity have attracted increasing interest owing to their advantages in cost and the convenience of oral administration 41. The multidomain architecture and mechanisms of action of clostridial toxins have provided distinct options for blocking their toxicity. Indeed, previous efforts have identified diverse inhibitors that act on distinct steps of intoxication, such as the inhibition of toxin binding to cell receptors, the suppression of endosomal maturation, and the inactivation of GTD or CPD activities 23, 48, 49, 50, 51. Although which approach is the best remains unknown, we believe that an inhibitor that targets multiple steps of intoxication would be superior to those that act on a single step. In this study, we identified a series of caffeic acid derivatives that were potent inhibitors of TcdB via a phenotypic screen. Further mechanistic study revealed that CAPE, the phenethyl alcohol ester of caffeic acid, could block InsP6-induced autoproteolysis of TcdB and inactivate its GTD function (Figure 10). Moreover, CAPE treatment significantly protected against CDI in a mouse model of infection. Our results suggested that caffeic acid derivatives, particularly CAPE, are promising lead compounds for the development of anti-virulence drugs that target TcdB. CAPE is a major component of propolis, a natural bee-produced substance that has been used as a remedy for thousands of years in Europe and Asia 52. As a result, further investigations on the therapeutic potential of propolis against CDI, either in combination with standard therapy or as a single therapy, would be of great interest.

A model for CAPE-mediated suppression of TcdB cytotoxicity and protection of mice from CDI.

The intoxication of cells by TcdB involves multistep mechanisms. Direct interaction of CAPE with full-length TcdB blocks InsP6-induced autoproteolysis, resulting in decreased production of active GTD. The binding between GTD and CAPE inhibits its glucosyltransferase activity, thus decreasing the glucosylation level of Rac1. CAPE treatment of CDI mice markedly restored the diversity and composition of the gut microbiota and induced changes in gut metabolites, which might have contributed to the therapeutic outcomes of CDI following CAPE treatment.

Antibiotic-induced gut microbiota dysbiosis is a major risk factor for CDI 53. Thus, reconstituting the gut microbiota is a promising strategy for CDI therapy. Fecal microbiota transplantation (FMT) is considered an effective CDI therapeutic strategy and is widely used in clinical practice 54. Despite the high cure rate and increasing popularity of FMT, patients remains limited acceptance. Another prominent advantage of anti-virulence drugs is that they allow maintenance of the normal microbiota, which is of critical importance for the treatment of CDI. Importantly, CAPE treatment of C. difficile-infected mice restored the structure and composition of the microbiota (Figure 10). In particular, the abundance of Bacteroides, a critical gut microbiota component that is required for resistance to C. difficile colonization 26, 55, 56, 57, was significantly elevated upon CAPE treatment. Interestingly, Bacteroides were also predominant in CDI patients administered RBX2660 (Rebyota), an FDA-approved fecal microbiota product 58, 59. The increased diversity of the gut microbiota might contribute to the therapeutic effects of CAPE in CDI.

In addition to the microbiota, gut metabolites also play crucial roles in the pathogenesis of CDI, either by promoting or suppressing disease progression30. For example, bile acids and amino acids promote spore germination and pathogenesis, while secondary bile acid and short-chain fatty acids aid in colonization resistance and suppress spore germination 30. In addition to restoration of the gut microbiota, metabolomics analysis of mouse fecal samples revealed dramatic alterations in gut metabolites upon CAPE treatment (Figure 10). The differentially abundant metabolites were significantly enriched in the purine metabolism and bile metabolism pathways. Remarkably, the level of adenosine, an endogenous purine nucleoside that can regulate immune responses and inflammatory processes by binding to its receptors, such as its ability to inhibit the release of certain inflammatory mediators, was dramatically increased in CAPE-treated mice 60, 61. In contrast, the amount of uric acid, the final product of purine metabolism, was reduced. Previous studies showed that the increase in adenosine concentrations caused by the inhibition of adenosine deaminase via EHNA prevents C. difficile TcdA-mediated damage and inflammation 33. This protection might be a result of the anti-inflammatory effects of adenosine, which rely on the activation of its receptors 62. In addition, we observed a reduction in the amount of D-proline, which is one of the preferred amino acids utilized by C. difficile to generate energy through Stickland reactions 35. Melatonin is a ubiquitous indolamine that has broad biological activities 63. Notably, we detected a significantly increase in the amount of melatonin in the feces of mice that were administered CAPE. Importantly, an earlier study demonstrated a decreased risk of recurrent CDI in patients exposed to melatonin, possibly owing to its antibacterial and anti-inflammatory activities 37. In this study, we found that oral administration of exogenous melatonin alleviated the symptoms of CDI in a mouse model of infection. Although determining the exact mechanism will require further investigation, we propose that the melatonin-mediated inhibition of TcdB intoxication observed in this study might contribute to its therapeutic effect. Taken together, these results indicate that the alterations in gut metabolites caused by CAPE might also contribute to treatment outcomes for CDI patients.

Methods Cell lines

All cell lines (HeLa, Caco-2, Vero, HEK293T and THP-1) were obtained from ATCC (Manassas, VA). HeLa, Caco-2, Vero and HEK293T cells were cultured in Dulbecco’s modified Eagle’s medium (DMEM) (Invitrogen, Carlsbad, CA) supplemented with 10% fetal bovine serum (FBS) (Invitrogen) and 1% (w/v) penicillin-streptomycin. THP-1 cells were cultured in Roswell Park Memorial Institute (RPMI) 1640 medium (Invitrogen) supplemented with 10% FBS and 1% (w/v) penicillin-streptomycin. All cell lines were cultured at 37 °C in a 5% CO2 atmosphere.

Bacterial strains and reagents

C. difficile BAA-1870 (RT027/ST1) was kindly provided by Aiwu Wu (Guangzhou Medical University) and cultured in brain-heart infusion broth (BHI) supplemented with 5 g/L yeast extract and 0.1% (w/v) L-cysteine (BHIS). The compounds were identified from two main sources: FDA-approved drugs obtained from ApexBio Technology (Houston, TX, USA) and small molecule natural compounds obtained from Chengdu Ruifensi Biotechnology Co., Ltd. (Chengdu, China). LysoTracker Red DND-99, kanamycin, and cefoperazone sodium were purchased from Dalian Meilun Biotechnology Co., Ltd. (Dalian, China).

Mice

Male C57BL/6 mice aged 6-8 weeks were purchased from Liaoning Changsheng Biotechnology Co., Ltd., and housed in a specific pathogen-free (SPF) facility. The mice lived in a controlled environment with a temperature of 23±2 °C and a humidity of 55±10 °C. Prior to infection, the mice were acclimated to the environment for a minimum of one week. All mouse experiments were approved by the Jilin University Institutional Animal Care Committee, Jilin University and strictly conducted according to the guidelines of this committee (SY202402301).

Protein expression and purification

The expression and purification of full-length TcdB were conducted following the method described by Yang et al. 64. In brief, Bacillus megaterium harboring pHis1522-TcdB was inoculated into LB media supplemented with a final concentration of 10 μg/mL tetracycline. Once the OD600nm reached approximately 0.3, xylose (0.5% w/v) was added to induce protein expression at 37 °C for 16 h. Then, 100 mL of the bacterial cells was harvested by centrifugation, resuspended in 5 mL of lysis buffer containing 20 mM imidazole, 300 mM NaCl, 20 mM NaH2PO4, 500 μM EDTA, and protease inhibitor cocktail (adjusted to pH 8.0), and subjected to sonication. The cell lysate was then centrifuged at 13,400 × g for 20 min to remove cell debris. The TcdB protein was purified using nickel column affinity chromatography. The purity of the protein was confirmed by SDS-PAGE, and the protein was subsequently dialyzed against buffer containing 10% glycerol, 20 mM Tris (pH 7.5), and 150 mM NaCl. After the protein concentration was determined, the purified TcdB protein was aliquoted and stored at –80 °C for further use.

For the purification of GTD, amplified DNA products were inserted into pET28a (Novagen), and the Escherichia coli (E. coli) strain BL21 (DE3) was used for the expression and purification of GTD. Transformed E. coli pET28a-GTD was cultured in Luria–Bertani (LB) medium supplemented with 10 μg/mL kanamycin. Once the OD600 reached 0.6, the expression of proteins was induced with 0.2 mM isopropyl-β-d-thiogalactoside (IPTG) overnight at 18 °C. The bacterial cells were then harvested by centrifugation and subjected to sonication. The cell lysate was then centrifuged at 13,400 × g for 20 min to remove cell debris. GTD protein was purified using nickel affinity chromatography and dialyzed in buffer containing 10% glycerol, 20 mM Tris (pH 7.5), and 150 mM NaCl. The GTD mutants were introduced into pET28a-GTD using the Quick Change Site-Directed Mutagenesis Kit (Stratagene, San Diego, CA, USA) with the primers listed in Table S2 The subsequent purification was performed using the same procedure as described for wild-type GTD.

Reverse transcription-quantitative PCR

Total RNA was extracted by TRIzol and then reverse-transcribed into complementary DNA by a NovoScript® Plus All-in-one 1st Strand cDNA Synthesis SuperMix (E047; Novoprotein, China). mRNA expression of tcdB was measured by quantitative PCR with NovoScript®SYBR qPCR SuperMix Plus (E096; Novoprotein, China) using the primers shown in Table S3. 16S rRNA was used as internal control. The relative mRNA expression levels were calculated by the delta-delta CT method.

Growth curve assays

Growth curve assay was performed to estimate the influence of CAPE on the growth of C. difficile BAA-1870. The overnight culture of BAA-1870 was transferred to fresh BHIS medium and incubated at 37°C in an anaerobic atmosphere. The culture was supplemented with different concentrations of CAPE ranging from 0 to 64 μg/mL and the optical density at 600 nm was monitored at 4 h intervals to assess growth of BAA-1870.

Inhibition of spore outgrowth

The impact of CAPE on the sporulation of C. difficile BAA-1870 was assessed using the protocol described by Joshua et al 65. In brief, BAA-1870 cultures were treated with 8 or 32 μg/mL of CAPE for 48 h. The treated samples were then harvested, subjected to a double wash with PBS, and incubated in a 60°C water bath for 30 min. Subsequently, 20 μL of each spore suspension was plated onto BHIS agar supplemented with 0.1% taurine to facilitate germination. These plates were incubated anaerobically at 37°C for 5 d to allow for spore germination. Following this period, the number of resulting colonies on the plates was enumerated.

Cell rounding assay

Cell rounding assay was performed according to Kristina et al, with appropriate adjustments 48. Vero or HeLa cells were seeded in 96-well sterile plates using serum-free DMEM overnight, and the compounds were preincubated with TcdB for 1 h. After incubation, the TcdB-compound complex was added to cells for 1 h at 37 °C. The cell morphology was visualized under a light microscope.

LysoTracker assay

The pH of endosomes was determined by LysoTracker Red DND-99 49. Caco-2 or THP-1 cells in complete DMEM were seeded in 96-well sterile plates overnight. The media was changed to serum-free DMEM for 1 h, and CAPE was preincubated with TcdB for 1 h, followed by addition to the cells and incubation for 2 h. Baf-A1 served as a control. After incubation, LysoTracker and Hoechst dyes were added and the cells were incubated for 1 h. Excess dye was removed by changing the media, and alterations in lysosomal red intensity were documented using a fluorescence microscope.

In vitro TcdB autoprocessing assay

The self-cleavage ability of TcdB primarily depends on its cysteine protease activity 48. CAPE (1-16 μg/mL) was preincubated with TcdB for 1 h, followed by the addition of InsP6 and incubation at 37 °C for 6 h. Subsequently, loading buffer was added for SDS-PAGE analysis, and the results were visualized using Coomassie Brilliant Blue staining.

Rac1 glucosylation assay

Rac1 glucosylation assay was performed following the method described by John et al 50. HEK293T cells were seeded into a 6-well plate. The next day, TcdB and different concentrations of CAPE (4-32 μg/mL) were added to cells. After 2 h, the cells were harvested, lysed using lysis buffer, and subjected to centrifugation at 13,400 × g for 10 min to collect the supernatant. The collected supernatant was then mixed with loading buffer and subjected to SDS-PAGE followed by western blot analysis. The glycosylation level of Rac1 was evaluated by determining the levels of non-glycosylated Rac1 (BD Biosciences, Mississauga, ON) and total Rac1 (Abcam).

UDP-GloTM UDP-glucose hydrolase assay

GTD activity was assessed according to the manufacturer’s instructions 50. A total volume of 16 μL, consisting of 100 nM GTD, various concentrations of CAPE and glucosylation buffer, was prepared. Four microliters of UDP-glucose at a final concentration of 25 μM was added to the mixture, which was then allowed to react at room temperature for 15 min. To stop the reaction, 10 μL of the reaction mixture was removed and placed in a white polystyrene 96-well plate containing 10 μL of UDP detection reagent. The reaction took place at room temperature for 1 h. The luminescence value was measured at a 750 ms integral, and the GTD activity was determined based on these measurements. Circular dichroism (CD) spectroscopy

CD spectra of TcdB, TcdB-CAPE, GTD, and GTD-CAPE were obtained using a CD spectrophotometer (MOS-500; Bio-Logic, France). TcdB or GTD at a concentration of 0.2 mg/mL was incubated with 32 μg/mL CAPE. Changes in the secondary structure of TcdB or GTD were assessed at room temperature using a quartz cuvette with an optical path length of 1 mm. The scan wavelength range was set from 190 to 250 nm with a resolution of 0.2 nm and a bandwidth of 1 nm. The secondary structures of TcdB, TcdB-CAPE, GTD, and GTD-CAPE were analyzed using the BeStSel web server.

Molecular docking

CAPE was obtained as a ligand from PubChem, and its topology was generated using AmeberTools. The crystal structure of GTD was retrieved from the Protein Data Bank (PDB) and processed using AutoDockTools. Docking calculations were performed using AutodockVina, and molecular dynamics simulations were carried out using Gromacs 2020.6. The RMSD of the receptor and ligand was monitored to assess conformational changes during the binding process. The interaction energy was analyzed using the Poisson-Boltzmann surface area (MM-PBSA) method, which is a molecular mechanics approach.

Flow cytometry

Flow cytometry was used to quantify the cellular uptake of TcdB via endocytosis. To facilitate detection, full-length TcdB was labeled with Alexa Fluor®488 NHS Ester (Yeasen Biotechnology Co., Ltd., Shanghai, China). Prior to cell treatment, Alexa Fluor 488-TcdB was preincubated with CAPE at a concentration of 32 μg/mL for 30 min. The treated solution was then added to Caco-2 cells and incubated for 1 h at 37 °C. Following three washes with phosphate-buffered saline (PBS), the Caco-2 cells were collected, and the fluorescence signal was measured using flow cytometry.

Cytotoxicity assay

Vero or THP-1 cells were plated into 96-well plates. The next day, different concentrations of CAPE were prepared by diluting them in DMEM and adding them to the cells. After incubation at 37 °C for 3 h, the viability of cells was assessed using a Cell Counting Kit-8 (CCK-8) (Beyotime Biotechnology Co., Shanghai, China) according to the manufacturer’s instructions.

Biacore analysis

The affinity analysis of TcdB or GTD with CAPE was conducted using a Biacore T100 device (Biacore Inc., Uppsala, Sweden) according to the manufacturer’s protocol. A CM5 chip (General Electric Company, GE) was utilized to immobilize the TcdB or GTD proteins. Various concentrations of CAPE were injected, and all steps were performed in running buffer consisting of 5% DMSO in PBS-P20 (GE). Evaluation version 3.1 software provided by Biacore and the 1:1 Langmuir combined model were employed for data analysis.

Mouse infection model

For the acute toxicity model, after pretreatment of TcdB with CAPE (2 or 4 μg/mL) for 1 h, the mice were intraperitoneally injected with a complex of TcdB-solvent or TcdB-CAPE (30 ng of TcdB per mouse). The survival of mice was observed for 72 h.

For the CDI model, mice were given cefoperazone (0.5 g/L) in their drinking water for 5 days (from day –7 to day –2), followed by switching to sterilized water for 2 days. On day 0, the mice were challenged with 1 × 108 colony-forming units (CFUs) of C. difficile BAA-1870 (Ribotype 027) by oral gavage. For the CAPE or melatonin treatment group, mice were treated with CAPE (30 mg/kg by oral gavage) or melatonin (200 mg/kg by oral gavage) every day after C. difficile challenge. The weight loss and fecal state of the mice were monitored daily. Fecal samples were collected on days 0-3 to evaluate the stool water content as a marker of diarrhea. Additionally, fecal samples on day 3 were collected to assess C. difficile colonization. Colonic tissues were collected on day 3 post infection, stained with hematoxylin and eosin (H&E) and scored as previously described 66.

Fecal samples were collected on day 3 post infection, frozen in liquid nitrogen and transferred to the laboratory with dry ice for 16S rRNA sequencing and untargeted metabolomics analysis (Novogene Bioinformatics Technology Co., Ltd., Tianjin, China). The melatonin level in mouse stool supernatant was determined using the Mouse MT (Melatonin) ELISA Kit (Sangon Biotechnology Co., Ltd., Shanghai, China).

Statistical analysis

Western blot densitometry and LysoTracker fluorescence intensity were measured using ImageJ software (NIH, Bethesda, MD, USA). Statistical analysis was performed using GraphPad Prism 9.0 (GraphPad Software, USA). Unpaired two-tailed Student’s t tests were used for comparing two groups, and one-way analysis of variance (ANOVA) was used for analyzing data from more than three groups. The statistical significance was determined using the log-rank (Mantel–Cox) test (survival rates) and Mann–Whitney U test (tissue bacterial load) in vivo. All the data are expressed as the mean ± standard deviation (SD). * indicates p <0.05 and ** indicates p <0.01.

eLife Assessment

This valuable study by Guo and colleagues reports the inhibitory activity of caffeic acid phenethyl ester (CAPE) against TcdB, a key toxin produced by Clostridioides difficile. C. difficile infections are a major public health concern, and this manuscript provides interesting data on toxin inhibition by CAPE, a potentially promising therapeutic alternative for this disease. The strength of the evidence to support the conclusions is solid, with some concerns about the moderate effects on the mouse infection model and direct binding assays of CAPE to the toxin.

Reviewer #1 (Public review):

Summary:

In this manuscript, Guo and colleagues used a cell rounding assay to screen a library of compounds for inhibition of TcdB, an important toxin produced by Clostridioides difficile. Caffeic acid and derivatives were identified as promising leads, and caffeic acid phenethyl ester (CAPE) was further investigated.

Strengths:

Considering the high morbidity rate associated with C. difficile infections (CDI), this manuscript presents valuable research in the investigation of novel therapeutics to combat this pressing issue. Given the rising antibiotic resistance in CDI, the significance of this work is particularly noteworthy. The authors employed a robust set of methods and confirmatory tests, which strengthen the validity of the findings. The explanations provided are clear, and the scientific rationale behind the results is well-articulated. The manuscript is extremely well written and organized. There is a clear flow in the description of the experiments performed. Also, the authors have investigated the effects of CAPE on TcdB in careful detail, and reported compelling evidence that this is a meaningful and potentially useful metabolite for further studies.

Weaknesses:

The authors have made some changes in the revised version. However, many of the changes were superficial, and some concerns still need to be addressed. Important details are still missing from the description of some experiments. Authors should carefully revise the manuscript to ascertain that all details that could affect interpretation of their results are presented clearly. For instance, authors still need to include details of how the metabolomics analyses were performed. Just stating that samples were "frozen for metabolomics analyses" is not enough. Was this mass-spec or NMR-based metabolomics. Assuming it was mass-spec, what kind? How was metabolite identity assigned, etc? These are important details, which need to be included. Even in cases where additional information was included, the authors did not discuss how the specific way in which certain experiments were performed could affect interpretation of their results. One example is the potential for compound carryover in their experiments. Another important one is the fact that CAPE affects bacterial growth and sporulation. Therefore, it is critical that authors acknowledge that they cannot discard the possibility that other factors besides compound interactions with the toxin are involved in their phenotypes. As stated previously, authors should also be careful when drawing conclusions from the analysis of microbiota composition data, and changes to the manuscript should be made to reflect this. Ascribing causality to correlational relationships is a recurring issue in the microbiome field. Again, I suggest authors carefully revise the manuscript and tone down some statements about the impact of CAPE treatment on the gut microbiota.

Reviewer #2 (Public review):

I appreciate the author's responses to my original review. This is a comprehensive analysis of CAPE on C. difficile activity. It seems like this compound affects all aspects of C. difficile, which could make it effective during infection but also make it difficult to understand the mechanism. Even considering the authors responses, I think it is critical for the authors to work on the conclusions regarding the infection model. There is some protection from disease by CAPE but some parameters are not substantially changed. For instance, weight loss is not significantly different in the C. difficile only group versus the C. difficile + CAPE group. Histology analysis still shows a substantial amount of pathology in the C. difficile + CAPE group. This should be discussed more thoroughly using precise language.

Reviewer #3 (Public review):

Summary:

The study is well written, and the results are solid and well demonstrated. It shows a field that can be explored for the treatment of CDI

Strengths:

Results are really good, and the CAPE shows a good and promising alternative for treating CDI.

Weaknesses:

Some references are too old or missing.

Comments on revisions:

I have read your study after comments made by all referees, and I noticed that all questions and suggestions addressed to the authors were answered and well explained. Some of the minor and major issues related to the article were also solved. I am satisfied with all the effort given by the authors to improve their manuscript.

Author response:

The following is the authors’ response to the previous reviews

Public Reviews:

Reviewer #1 (Public review):

Summary:

In this manuscript, Guo and colleagues used a cell rounding assay to screen a library of compounds for inhibition of TcdB, an important toxin produced by Clostridioides difficile. Caffeic acid and derivatives were identified as promising leads, and caffeic acid phenethyl ester (CAPE) was further investigated.

Strengths:

Considering the high morbidity rate associated with C. difficile infections (CDI), this manuscript presents valuable research in the investigation of novel therapeutics to combat this pressing issue. Given the rising antibiotic resistance in CDI, the significance of this work is particularly noteworthy. The authors employed a robust set of methods and confirmatory tests, which strengthen the validity of the findings. The explanations provided are clear, and the scientific rationale behind the results is well-articulated. The manuscript is extremely well written and organized. There is a clear flow in the description of the experiments performed. Also, the authors have investigated the effects of CAPE on TcdB in careful detail, and reported compelling evidence that this is a meaningful and potentially useful metabolite for further studies.

Weaknesses:

The authors have made some changes in the revised version. However, many of the changes were superficial, and some concerns still need to be addressed. Important details are still missing from the description of some experiments. Authors should carefully revise the manuscript to ascertain that all details that could affect interpretation of their results are presented clearly. For instance, authors still need to include details of how the metabolomics analyses were performed. Just stating that samples were "frozen for metabolomics analyses" is not enough. Was this mass-spec or NMR-based metabolomics. Assuming it was mass-spec, what kind? How was metabolite identity assigned, etc? These are important details, which need to be included. Even in cases where additional information was included, the authors did not discuss how the specific way in which certain experiments were performed could affect interpretation of their results. One example is the potential for compound carryover in their experiments. Another important one is the fact that CAPE affects bacterial growth and sporulation. Therefore, it is critical that authors acknowledge that they cannot discard the possibility that other factors besides compound interactions with the toxin are involved in their phenotypes. As stated previously, authors should also be careful when drawing conclusions from the analysis of microbiota composition data, and changes to the manuscript should be made to reflect this. Ascribing causality to correlational relationships is a recurring issue in the microbiome field. Again, I suggest authors carefully revise the manuscript and tone down some statements about the impact of CAPE treatment on the gut microbiota.

Thanks for your constructive suggestion. We have carefully revised the manuscript according to your suggestions.

Reviewer #2 (Public review):

I appreciate the author's responses to my original review. This is a comprehensive analysis of CAPE on C. difficile activity. It seems like this compound affects all aspects of C. difficile, which could make it effective during infection but also make it difficult to understand the mechanism. Even considering the authors responses, I think it is critical for the authors to work on the conclusions regarding the infection model. There is some protection from disease by CAPE but some parameters are not substantially changed. For instance, weight loss is not significantly different in the C. difficile only group versus the C. difficile + CAPE group. Histology analysis still shows a substantial amount of pathology in the C. difficile + CAPE group. This should be discussed more thoroughly using precise language.

Thanks for your constructive suggestion. We have carefully revised the manuscript according to your suggestions.

Reviewer #3 (Public review):

Summary:

The study is well written, and the results are solid and well demonstrated. It shows a field that can be explored for the treatment of CDI

Strengths:

Results are really good, and the CAPE shows a good and promising alternative for treating CDI.

Weaknesses:

Some references are too old or missing.

Comments on revisions:

I have read your study after comments made by all referees, and I noticed that all questions and suggestions addressed to the authors were answered and well explained. Some of the minor and major issues related to the article were also solved. I am satisfied with all the effort given by the authors to improve their manuscript.

Thanks again for your review.

Recommendations for the authors:

Reviewer #1 (Recommendations for the authors):

The legend of Figure 3SB is incorrect. It should read "Growth curves of C. difficile BAA-1870 in the presence of varying concentrations of CAPE (0-64 µg/mL)". Also, there is something wrong with the symbols in this figure. I suspect what is happening is that the symbols for the concentrations of 32 and 64 µg/mL are superimposing, but this is a problem because the lower line looks like a closed circle, which is supposed to represent the condition where no CAPE was added. The authors should change the symbols to allow clear distinction between each of the conditions.

Thanks for your constructive suggestion. We have modified the panel and figure legend in Figure 3SB. The concentrations of 32 μg/mL and 64 μg/mL are quite similar, which makes it challenging to differentiate between the corresponding data points on the graph. To enhance clarity, we have utilized distinct colors to help distinguish these closely valued lines as effectively as possible.

Since the authors observed a significant effect of CAPE on both bacterial growth and spore production, their discussion and conclusions need to reflect the fact that the effects observed can no longer be attributed solely to toxin inhibition.

Thanks for your comments. We have modified the corresponding description according to your suggestions.

In lines 43-45, authors state that "CAPE treatment of C. difficile-challenged mice induces a remarkable increase in the diversity and composition of the gut microbiota (e.g., Bacteroides spp.)". It is still unclear to this reviewer why mention Bacteroides between parentheses. Does this mean that there was an increase in the abundance of Bacteroides? If that is the case this needs to be stated more clearly.

Thanks for your comments. Treatment with CAPE indeed significantly increased the abundance of Bacteroides spp. in the gut microbiota (Figure 7H-J). However, to avoid ambiguity in the abstract, we have chosen to delete the specific mention of Bacteroides spp. within the parentheses.

The modifications made to lines 132-135 still do not address my concern. Authors stated in the manuscript that "compounds that were not bound to TcdB were removed". But how was this done? This needs to be clearly explained in the manuscript. In the response to reviewers document, authors state that this was done through centrifugation. But given that the goal here is to separate excess of small molecule from a protein target, just stating that centrifugation was used is not enough. Did the authors use ultracentrifugation? What were the conditions employed. This is critical so that the reader can assess the degree of compound carryover that may have occurred. Also, authors need to clearly acknowledge the caveats of their experimental design by stating that they cannot rule out the contribution of compound carryover to their results.

Thanks for your comments. We employed ultrafiltration centrifugal partition to remove the unbound small molecule compounds. Due to the large molecular weight of TcdB, approximately 270 kDa, we selected a 100 kDa molecular weight cutoff ultrafiltration membrane. The centrifugation was performed at 4000 g for 5 min to eliminate the compounds that did not bind to TcdB. We have incorporated the relevant methods and discussed the potential impacts on the respective sections of the manuscript.

In line 142, authors added the molar concentration of caffeic acid, as requested. Although this helps, it is even more important that molar concentrations are added every time a compound concentration is mentioned. For instance, just 2 lines down there is another mention of a compound concentration. It would be informative if authors also added molar concentrations here and throughout the manuscript.

Thanks for your comments. In our initial test design, we have utilized the concentration unit of μg/mL. However, during the conversion to μM using the dilution method, some values do not result in neat, whole numbers. For instance, the conversion of 32 μg/mL of caffeic acid phenyl ethyl ester yields 112.55 μM, which appears somewhat irregular when expressed in this manner.

Line 277. For the sake of clarity, I would strongly suggest that authors use the term "control mice" instead of "model mice".

Thanks for your comments. We have modified “model mice” to “control mice” throughout the manuscript.

In line 302, the word taxa should not be capitalized. I capitalized it in my original comments simply to draw attention to it.

Thanks for your comments. We have modified this word.

In the section starting in line 318, authors still need to include details of how the metabolomics analyses were performed. Just stating that samples were "frozen for metabolomics analyses" is not enough. Was this mass-spec or NMR-based metabolomics. Assuming it was mass-spec, what kind? How was metabolite identity assigned? Etc, etc. These are important details, which need to be included.

Thanks for your comments. We have added some metabolomics methods in the corresponding section.

In line 338, the authors misunderstood my original comment. This sentence should read "...the final product of purine degradation, were markedly decreased in mice after...".

Thanks for your comments. We have modified this sentence.

Panels of figure 3 are still incorrectly labeled. The secondary structure predictions are shown in A and C, not A and B as is currently stated in the legend.

Thanks for your comments. We have modified the figure legend in Figure 3.

About Figure 5C, I think the authors for the clarification, but this explanation should be included in the figure legend.

Thanks for your comments. We have added the relevant information to the figure legend.

Version 3

Abstract

Abstract

Clostridioides difficile infection (CDI) is the leading cause of hospital-acquired diarrhea that seriously threatens public health. The disruption of normal gut microbiota by the use of broad-spectrum antimicrobial agents enables C. difficile to proliferate in the colon. The emergence and prevalence of hypervirulent C. difficile strains result in increased morbidity, mortality, and high recurrence rates of CDI, thus creating a pressing need for novel therapeutics. The multi-domain toxins TcdA and TcdB are the primary determinants of CDI pathogenesis, rendering them ideal drug targets in the anti-virulence paradigm. In this study, we identified caffeic acid and its derivatives from natural compounds library as active inhibitors of TcdB via a cell-based high-throughput phenotypic screening. Further mechanistic investigations revealed that caffeic acid phenethyl ester (CAPE) could directly bind to TcdB, thus suppressing InsP6-induced autoproteolysis and inhibiting glucosyltransferase activity. CAPE treatment remarkably reduces the pathology of CDI in a murine infection model in terms of alleviated diarrhea symptoms, decreased bacterial colonization, and relieved histopathological lesions. Moreover, CAPE treatment of C. difficile-challenged mice induces a remarkable increase in the diversity and composition of the gut microbiota and alterations of gut metabolites (e.g., adenosine, D-proline, and melatonin), which might partially contribute to the therapeutic outcomes of CAPE against CDI. Our results reveal the potential of CAPE as a therapeutic for the management of CDI, or CAPE might serve as a lead compound for the development of antivirulence drugs targeting TcdB.

Introduction

Clostridioides difficile (C. difficile) is an obligate anaerobic, spore-forming, gram-positive bacterial pathogen that is the most common cause of hospital-acquired and antibiotic-associated gastrointestinal diseases, including mild diarrhea and life-threatening pseudomembranous colitis 1, 2, 3. In the United States, C. difficile is responsible for almost half a million infections and approximately 29,000 deaths per year, resulting in an estimated annual healthcare-associated cost of up to $4.8 billion 4, 5. Importantly, the incidence, severity, mortality and healthcare costs of C. difficile infection (CDI) have dramatically increased in the 21st century, at least partially owing to the frequent use of antimicrobial agents and the emergence of hypervirulent strains (e.g., PCR ribotypes 027 and 078) 6, 7, 8. Limited options are available for the treatment of CDI, while antibiotics including metronidazole, fidaxomicin and vancomycin, are the mainstays of clinical practice 9, 10. However, antibiotic treatment is ineffective in more than 35% of CDI patients and can lead infection recurrence 11, 12. Moreover, the rapid evolution of antibiotic resistance in C. difficile has further compromised the therapeutic efficacy of current treatment options 13. Therefore, the high morbidity and mortality, high recurrence rates, and emergence of multiple antibiotic-resistant isolates have made this pathogen a major threat to public health; these factors highlight the urgent need for novel therapies.

C. difficile widely exists in the environment and in the intestinal tract of humans and animals and is mainly transmitted via the fecal-oral route 14. The pathogenesis of CDI is correlated with the use of antibiotics, which impair the structure and composition of the normal gut microbiota, thus allowing C. difficile to proliferate in the colon 14. Therefore, previous antibiotic therapy is the major risk factor for the development of CDI. In the colon, the dormant spores of C. difficile germinate into oxygen-sensitive vegetative cells that produce toxins and cause symptoms of disease 15. C. difficile secretes two large clostridial exotoxins (TcdA and TcdB), which are directly responsible for the clinical pathology of CDI 16, 17. Epidemiological studies suggest that most disease-causing C. difficile strains have genes that encode TcdB 18, 19. TcdA and TcdB have identical functional domains: a N-terminal glucosyltransferase catalytic domain (GTD), a cysteine protease active domain (CPD), a transmembrane domain (DRBD), and a C-terminal receptor binding domain (CROPs) 20. When secreted into the colon, both toxins are internalized by host cells through receptor-mediated, clathrin-dependent endocytosis. Following a decrease in endosomal pH, the toxins undergo conformational changes in the DRBD, resulting in insertion of the toxin into the endosome membrane, pore formation, and the release of the GTD and CPD domains into the cytosol. Then, CPD binding with cytosolic inositol hexakisphosphate (InsP6) activates its protease activity, leading to the cleavage and release of the GTD. Eventually, the GTD utilizes UDP-glucose to glucosylate Rho and/or Ras family GTPases, thus disrupting GTPase signaling and resulting in deleterious effects, including cytopathic “rounding”, dysregulation of the actin cytoskeleton, apoptosis, and pyroptosis 1. Although both toxins are cytotoxic to host cells, TcdB shows greater toxicity than TcdA 21.

Disruption of the gut microbiota by antimicrobial agents results in high recurrence rates, which limits the application of antimicrobial therapy for CDI 14. Consequently, there is a critical need to develop nonantibiotic treatments for CDI, either as standalone or augmented therapies. Importantly, strategies targeting bacterial virulence factors, including toxins, are termed anti-virulence strategies and have attracted interest in the 21st century 22. Owing to its essential role in C. difficile pathogenesis, TcdB has become an ideal target for the development of anti-virulence drugs 23. In this study, caffeic acid and a series of its derivatives were found to inhibit TcdB-mediated cell rounding via cell-based screening. Among these potential inhibitors, caffeic acid phenethyl ester (CAPE) had the greatest inhibitory effect. Mechanistically, CAPE bound to both full-length TcdB and the GTD fragment, leading to alterations in the secondary structure. This interaction suppressed the InsP6-mediated self-cleavage of TcdB, as well as the enzymatic activity of GTD, thereby decreasing the level of glucosylated Rac1 in TcdB-treated host cells and further inhibiting the cytopathology and cytotoxicity of TcdB. Oral administration of CAPE significantly protected against CDI in models of animal infection. In addition to direct toxin inhibition, the therapeutic effect may also be attributed to CAPE-induced alteration of the gut microbiota (e.g., Bacteroides spp.) as well as gut metabolites (e.g., adenosine, D-proline, and melatonin). Overall, our study revealed that caffeic acid and its derivatives, particularly CAPE, may serve as promising lead compounds for the development of anti-virulence drugs to treat CDI.

Results

CAPE inhibits cell rounding induced by TcdB

TcdB can inactivate Rho GTPases via glucosylation, leading to the loss of cytoskeletal structure and cell-rounding. Consequently, we utilized this phenotype to identify potential TcdB inhibitors from 2076 chemicals, including 1515 FDA-approved drugs and 561 natural compounds. The minimal concentration of recombinant His6-TcdB (0.2 ng/mL) that induced 100% cell rounding within 1 h was used for drug screening. This concentration of TcdB was preincubated with 8 μg/mL of the individual compounds for 1 h, the compounds that were not bound to TcdB were removed by ultrafiltration centrifugal partition and then added to Vero cell cultures. Residual compounds that have not bound to TcdB may lead to an exaggerated assessment of their efficacy and could also result in misleading positive outcomes. As the effective concentration for cellular interactions may inadvertently be increased, such carryover could distort the determination of the compound’s minimum effective concentration. At 1 h post-TcdB treatment, cell morphology was visualized under a light microscope (Figure 1A). These analyses allowed us to identify 22 candidate compounds with potent inhibitory effects (Figure 1B). Interestingly, six out of the 22 compounds were caffeic acid derivatives. The natural compound library contained eight caffeic acid derivatives, of which methyl caffeic acid and ferulic acid displayed no efficacy. Caffeic acid itself only slightly inhibited TcdB-induced cell rounding, with a 50% inhibitory concentration (IC50) of 527.5 μg/mL (1.86 mM) (Table S1). Notably, except for rosmarinic acid, the IC50 values for the other screened caffeic acid derivatives were less than 6 μg/mL, suggesting that these compounds potently inhibited TcdB-mediated toxicity (Table S1). CAPE is the phenethyl alcohol ester of caffeic acid and a main active component of propolis with broad pharmacological properties, such as anticancer, anti-inflammatory, antiviral, antibacterial, antifungal, antioxidant, and immunomodulatory activities 24. As CAPE has the lowest IC50 (3.0 μg/mL) among the active caffeic acid derivatives, it was selected as the representative compound for subsequent investigations (Table S1). After preincubation with increasing concentrations of CAPE, dose-dependent inhibition of cell rounding was observed in both Vero and HeLa cells (Figure 1C). The protective influence of CAPE against TcdB-induced Vero cell death remained evident even when the treatment duration was prolonged to 12 h (Figure S1A-B).

CAPE suppresses TcdB-mediated cell rounding.

(A) Schematic of the process used for the identification of small molecules that inhibit the cell rounding phenotype induced by TcdB. TcdB (0.2 ng/mL) was preincubated with 8 μg/mL of individual compounds prior to addition to the cell cultures. (B) Active compounds were identified from the chemical library consisting of 2076 chemicals. The red dots represent caffeic acid and its derivatives. (C) Dose-dependent CAPE inhibition of TcdB-induced rounding of Vero (upper) and HeLa (lower) cells. Scale bar, 100 μm. The dose titration curves are presented on the right. (D) The survival rates of mice that were intraperitoneally injected with CAPE-pretreated or native TcdB (30 ng). * p <0.05 and ** p <0.01 by log-rank (Mantel-Cox) test.

In addition, we also determined whether inhibition of TcdB by CAPE could lead to loss of toxin function in vivo using a murine systemic intoxication model. A lethal dose of recombinant His6-TcdB (30 ng) was preincubated with either DMSO or CAPE (2 or 4 μg/mL) for 1 h at 37 °C, followed by intraperitoneal injection into mice. Notably, all mice in the DMSO group (n=10) died within 36 h after toxin injection. However, 100% of mice pretreated with CAPE 4 (μg/mL) toxin survived at the monitoring endpoint (72 h) (n=10) (Figure 1D). Furthermore, we evaluated the efficacy of CAPE following an intraperitoneal injection by administering it 2 h prior to the injection of TcdB, in order to gauge the survival rate of the mice. The outcomes revealed that a 2 h pretreatment with CAPE at doses of 30 and 50 mg/kg significantly enhanced the mice’s protection rate by 57.1% (Figure S2). Simultaneously, we assessed the impact of CAPE on C.difficile itself. CAPE markedly repressed tcdB gene expression at a concentration of 16 μg/mL (Figure S3A) and inhibited the growth and spore formation of C. difficile BAA-1870 at a concentration of 32 μg/mL (Figure S3B-C). Taken together, these data indicate that CAPE is a potent toxin inhibitor of TcdB and may serve as a promising antimicrobial agent.

CAPE decreases InsP6-induced autoproteolysis of TcdB and inhibits the enzymatic activity of GTD

TcdB is a multi-domain containing toxin that induces toxicity in host cells via an intricate and multistep mechanism. This mechanism involves (i) receptor-mediated endocytosis, (ii) pore formation and subsequent membrane translocation, (iii) autoproteolysis and release of GTD, and (iv) the glucosylation of small GTPases (Figure 2A). To determine the mechanism by which CAPE suppresses the cytotoxicity of TcdB, we first labeled full-length TcdB with Alexa Fluor 488; then, the concentration of the endocytosed toxin was determined via flow cytometry. Apparently, incubation of CAPE with TcdB did not significantly affect its entry into host cells (Figure 2B). Next, we stained CAPE-treated cells with LysoTracker, a fluorescent dye that accumulates in the acidic organelles of live cells. In contrast to the vacuolar-type H+-ATPase (V-ATPase) inhibitor bafilomycin A1 (Baf-A1), CAPE treatment did not alter the intensity of LysoTracker fluorescence staining in either THP-1 or Caco-2 cells, suggesting that inactivation of TcdB by CAPE is not due to changes in the pH of endosomal compartments (Figure 2C-D and Figure S4A-B).

Mechanism by which CAPE suppresses TcdB intoxication.

(A) Schematic illustration of the multistep mechanism of TcdB intoxication of host cells. (B) Determination of internalized TcdB levels. Alexa Fluor 488-labeled TcdB was incubated with 32 μg/mL CAPE for 30 min before it was added to the cell culture. The cellular uptake of TcdB was measured by flow cytometry analysis. (C) Lysosomal activity in host cells treated with CAPE was measured via LysoTracker staining. The nuclei were stained with DAPI. Bar, 100 μm. (D) The intensities of LysoTracker staining were quantified using ImageJ. * p <0.05; ** p <0.01 using one-way ANOVA. (E-F) CAPE inhibits InsP6-induced autoprocessing of full-length TcdB. Two micrograms of TcdB was preincubated with increasing concentrations of CAPE for 1 h. Self-cleavage of TcdB was initiated by the addition of 3 μg of InsP6 and the cells were incubated for 6 h at 37 °C. Reactions were terminated with 5x SDS loading buffer, and the release of GTD was visualized by CBB staining of the SDS-PAGE gel (E). The level of released GTD in panel E was quantified by ImageJ (F) * p <0.05; ** p <0.01 using one-way ANOVA. (G) Dose-dependent inhibition of the GTD activity of TcdB by CAPE. Recombinant GTD of TcdB was mixed with different concentrations of CAPE (0-32 μg/mL) and incubated for 15 min at 25 °C. GTD activity was measured by the UDP-Glo™ glycosyltransferase assay. (H-I) Intracellular Rac1 glucosylation in cells treated with TcdB. Native or CAPE-pretreated TcdB was added to the cell culture and cells were incubated for 2 h. The cells were lysed, and the glucosylation of Rac1 was analyzed by western blotting using antibodies recognizing total Rac1 and unglycosylated Rac1 (H). The level of Rac1 glucosylation shown in H was quantified by ImageJ (I). * p <0.05; ** p <0.01 using one-way ANOVA.

These observations indicate that CAPE might exert its effects after membrane translocation. Indeed, preincubation of full-length TcdB with CAPE blocked the InsP6-induced autoproteolysis of TcdB, as revealed by the gradual decrease in the production of released GTP, followed by the addition of increasing amounts of CAPE in vitro (Figure 2E-F). Furthermore, we determined the glucosyltransferase activity using recombinant GTD in the absence or presence of different doses of CAPE. Strikingly, incubation with 2 μg/mL CAPE resulted in a significant reduction in GTD activity, whereas 32 μg/mL CAPE compromised 87.5% of the enzymatic activity (Figure 2G). Importantly, in response to suppression of InsP6-induced autoproteolysis and GTD function, treatment of TcdB-challenged cells with CAPE markedly decreased the level of glucosylated Rac1 in a dose-dependent manner (Figure 2H-I). Taken together, these results suggest that CAPE disrupts InsP6-induced autoproteolysis of TcdB and inhibits the enzymatic activity of GTD, thereby decreasing the glucosylation of Rac1 and alleviating cell rounding.

CAPE alters the secondary structure of full-length TcdB and GTD

To further clarify the mechanisms used by CAPE to inactivate TcdB, we determined the secondary structures of full-length TcdB and GTD in the absence or presence of CAPE by circular dichroism spectroscopy. Apparently, incubation of CAPE with full-length TcdB resulted in a significant increase in the α-helix content (15.4% to 21.0%) and a concomitant decrease in the β-pleated sheet content (34.5% to 20.9%) (Figure 3A-B). Similarly, the secondary structure of GTD was markedly altered upon incubation with CAPE, with the proportion of α-helices decreasing from 37.9% to 0.0% and the proportion of β-pleated sheets decreasing from 37.2% to 10.9% (Figure 3C-D). Taken together, these data suggest that CAPE may interact with full-length TcdB and GTD to induce conformational changes, thus impairing their functions.

CAPE alters the secondary structure of TcdB and GTD.

(A) The secondary structure of TcdB in the presence or absence of 32 μg/mL CAPE was determined by CD spectroscopy. (B) Calculated CD spectra of CAPE-treated TcdB. (C) The secondary structure of GTD in the presence or absence of 32 μg/mL CAPE was determined by CD spectroscopy. (D) Calculated CD spectra of CAPE-treated GTD. The wavelength for CD spectroscopy was set at 190-250 nm.

CAPE directly interacts with full-length TcdB and GTD

To investigate the potential interaction between CAPE and TcdB or GTD, the binding constants were determined by surface plasmon resonance (SPR) analysis. The results showed that CAPE has strong binding affinity with full-length TcdB and GTD, and the KD values were 1.269×10-5 M and 1.007×10-3 M, respectively (Figure 4A-B). The interaction between CAPE and TcdB exhibits a slow approach to equilibrium, which may primarily be attributed to the rapid occupation of the protein’s primary binding sites by the small molecule during the initial phase. Following this, CAPE proceeds to bind to sites with secondary or reduced affinity, a process that prolongs the time required to achieve equilibrium. Furthermore, the possibility that CAPE binds to multiple sites on TcdB necessitates a duration for the exploration and occupation of these various sites before equilibrium is established. These data demonstrate the direct interaction between CAPE and full-length TcdB/GTD.

Direct interaction of CAPE with TcdB and GTD.

(A-B) Full-length TcdB or GTD was immobilized on a CM5 chip, followed by the injection of various concentrations of CAPE. The response unit (RU) value and affinities of CAPE interactions with TcdB (A) and GTD (B) were determined by a Biacore T100.

Next, we performed molecular docking and molecular simulation dynamics analyses to investigate the interaction between CAPE and GTD (Figure 5A). The results revealed that Leu265, Asp286, Thr465, Ile466, Leu519, and Trp520 in GTD were potential binding sites for CAPE (Figure 5A). The root mean square deviation (RMSD) of the GTD/CAPE complex was stable over a 100 ns period of simulation (Figure 5B). Further energy decomposition analysis showed that the side chains of Leu265, Thr465, Leu519, and Trp520 of GTD contributed to binding with CAPE via van der Waals interactions, among which Trp520 had the highest binding energy (Figure 5C). Moreover, we mutated each of the residues that were involved in binding and subsequently determined their glucosyltransferase activities in the presence or absence of CAPE. The enzymatic activities of these mutants were similar to that of wild-type GTD (Figure 5D). Mutations in GTDL265A, GTDT465A, GTDI466A, and GTDL519A did not affect the inhibitory effect of CAPE (Figure 5E-I). However, the inhibition of GTDW520A and GTDD286A by CAPE was notably compromised (Figure 5J-K).

Asp286 and Trp520 in the GTD are critical for CAPE-mediated inhibition of GTD.

(A) Three-dimensional structure determination of GTD with the CAPE complex was performed by molecular modeling. (B) The RMSD values of the GTD-CAPE complex. (C) Decomposition of the binding energy on a per-residue basis in the binding sites of the GTD-CAPE complex. Blue denotes the total binding energy, red signifies the electrostatic interactions, green corresponds to the van der Waals forces, and orange indicates solvation or hydration effects. The horizontal axis represents the mutation of the amino acid residue at the respective position to alanine. (D) Glucosyltransferase activity of wild-type GTD and the GTD mutants. (E-K) Residual activity of wild-type GTD (E) and its mutants GTDL265A (F), GTDT465A (G), GTDI466A (H), GTDL519A (I), GTDW520A (J) and GTDD286A (K) in the presence of increasing concentrations of CAPE (0-4 μg/mL). * p <0.05; ** p <0.01 using one-way ANOVA.

CAPE treatment reduces the pathology of CDI in a murine model

The abovementioned observation that CAPE suppresses the function of TcdB prompted us to further evaluate the potential therapeutic effects of this compound in CDI. To this end, we employed a well-established murine model in which infected mice can develop symptoms resembling those observed in human CDIs (25). First, cytotoxicity measurements revealed that the viability of both THP-1 and Vero cells was not affected by CAPE treatment (Figure S5A-B), thus indicating low toxicity. Mice were preconditioned with drinking water containing 0.5 g/L cefoperazone for 5 days, followed by normal drinking water for 2 days. After challenge with C. difficile BAA-1870 via gavage, mice were then treated daily with either vehicle or 30 mg/kg CAPE by gavage throughout the disease course (day 0 to day 5) (Figure 6A). Severe weight loss in C. difficile-infected mice was observed on days 2 and 3 post infection and was accompanied by diarrhea (Figure 6B). Notably, compared to the infection control group, CAPE treatment significantly increased the dry/wet weight ratio of the mouse feces, suggesting the alleviation of diarrhea symptoms (Figure 6C). Correspondingly, the weight loss of mice was relieved upon receiving CAPE (Figure 6B). Moreover, fecal pellets collected on day 3 were used to determine the bacterial burden. The results revealed that CAPE treatment significantly decreased the number of C. difficile CFUs (Figure 6D). On day 3 post-C. difficile infection, the colon was collected for histopathological analysis. Remarkably, C. difficile-infected mice exhibited conspicuous colon edema, epithelial cell injury and inflammatory factor infiltration (Figure 6E). In contrast, these histopathological lesions were significantly reduced in mice given CAPE, but pathological changes were still present (Figure 6F). This indicates that CAPE may have a limited capacity to reverse the histological damage induced by C. difficile. The residual pathology could be attributed to the complex interplay between the pathogen, the host immune response, and the tissue repair process, which may not be fully mitigated by the inhibitory effects of CAPE on TcdB.

CAPE treatment blocks the pathology of CDI in mice.

(A) Schematic of the process used for the establishment of the CDI mouse model. (B) Weight changes in CDI mice after CAPE or vehicle treatment (n=10). (C) The fecal dry/wet weight ratio in CDI mice treated with CAPE or vehicle (n=6). * p <0.05; ** p <0.01 using one-way ANOVA. (D) Bacterial colonization in the feces of CDI mice treated with CAPE or vehicle (n=15). * p <0.05; ** p <0.01 using the Mann Whitney U test. (E) Pathological analysis of colon sections from CDI mice treated with CAPE or vehicle. The black boxes are the regions that were magnified and are shown underneath. (F) Histopathological scores of vehicle-treated and CAPE-treated mice in terms of cell infiltration (orange), epithelial damage (purple), and edema (green). * p <0.05; ** p <0.01 using one-way ANOVA.

CAPE induces distinct microbiota changes in CAPE-treated mice when compared with control mice

Antibiotic-induced dysbiosis of the gut microbiota promotes the development of CDI 25. Therefore, we next determined whether CAPE has any impact on gut microbial communities that may contribute to its therapeutic efficacy. To this end, on the 3rd day after C. difficile challenge, fecal samples from the control group and CAPE treatment group were collected and subjected to microbiota analysis via 16S rRNA gene sequencing. Remarkably, as reflected by the Venn diagram, CAPE treatment resulted in significant changes in the microbial composition compared to those in the control group, and only 16.42% of the operational taxonomic units (OTUs) were shared (Figure 7A). In addition, both the α- and β-diversity of the gut microbiota in CAPE-treated mice were significantly greater than those in control mice. Specifically, the increase in the Chao1 richness estimator of mice in the CAPE treatment group suggested a greater number of microbial species and increased evenness in the fecal samples (Figure 7B-C); the increase in the Shannon-Wiener diversity index and the Simpson diversity index reflected a significant increase in microbiota diversity (Figure 7D-E). Principal coordinate analysis (PCoA) revealed distinct confidence ellipses between the bacterial communities of the CAPE-treated group and the control group, thereby revealing substantial differences in their microbiota profiles (Figure 7F).

The influence of CAPE treatment on the diversity and composition of the gut microbiota in CDI mice.

(A) Venn diagram illustrating the numbers of differentially abundant bacteria identified using a single model and their cooperation. (B-E) Quantification of the α diversity of the gut microbiota according to the observed species richness (B-C), Shannon index (D), and Simpson index (E). (F) PCoA showing β diversity based on the weighted UniFrac distance. (G-H) Relative abundance of the top 10 predominant bacteria classified at the phylum (G) and genus (H) levels. (I) Significant changes in flora among the control and CAPE groups, as measured by LEfSe analysis (LDA score [log 10] > 3.5). (J-M) Relative abundances of Bacteroides (J), Muribaculum (K), Faecalibaculum (L), and Parasutterella (M) in mice in the control and CAPE treatment groups. * p <0.05; ** p <0.01 using an unpaired Student’s t test.

A comparative analysis was performed to examine changes at the phylum level following CAPE treatment. Bacteroidota, Firmicutes, and Proteobacteria were the major constituents in both the CAPE treatment and control groups, suggesting that the administration of CAPE did not significantly alter the mouse microbiota at the phylum level (Figure 7G). Next, we compared the 10 most abundant taxa at the genus level. Notably, CAPE treatment markedly elevated the abundance of Bacteroides and decreased the abundance of Muribaculaceae and Parabacteroides (Figure 7H). Leveraging linear discriminant analysis of effect sizes (LEfSe), the microbiota was examined at the genus level to identify bacteria exhibiting substantial differences in abundance between the two cohorts (LDA score > 3.5, p < 0.05). After CAPE treatment, the relative abundance of Bacteroides, Ruminococcus_gnavus_group, Muribaculum and Faecalibaculum OTUs notably increased, while that of Hungatella, Parasutterella and Muribaculaceae OTUs significantly decreased (Figure 7I). Bacteroides, Muribaculum and Faecalibaculum improve colonization resistance to C. difficile 26, 27, 28. Importantly, the abundance of these bacterial species increased following CAPE treatment (Figure 7J-L). In contrast, the abundance of Parasutterella, a bacterium that facilitates C. difficile colonization 29, was reduced in mice receiving CAPE (Figure 7M). Taken together, these findings suggest that the gut microbiota alteration might contribute modestly to the observed effects.

CAPE alters the levels of gut metabolites in C. difficile-infected mice

In addition to the crucial role of the gut microbiota in conferring colonization resistance against C. difficile, gut microbiota-derived metabolites significantly contribute to the disease process 30. Metabolomic analysis was performed to evaluate the effects of CAPE treatment on gut metabolites in mice. On the 3rd day after infection, the feces of mice in each group were frozen by liquid nitrogen for metabolomics analysis. Specifically, utilizing mass spectrometry-based metabolomics, a simple screening of each metabolite is conducted for parameters such as retention time and mass-to-charge ratio. Subsequently, peak area calibration is performed using the first quality control (QC) sample to enhance the accuracy of identification. The peak extraction is then conducted with settings for mass deviation at 5 ppm, signal intensity deviation at 30%, minimum signal intensity, and adduct ions. Simultaneously, quantification of peak areas is carried out, and target ions are integrated. Molecular formulas are predicted using molecular ion peaks and fragment ions, and these are compared with databases such as mzCloud, mzVault, and Masslist. Background ions are removed using blank samples. The raw quantification results are standardized using the formula: original quantification value of the sample / (total metabolite quantification value of the sample / total metabolite quantification value of QC1 sample), to obtain relative peak areas. Compounds with a coefficient of variation (CV) greater than 30% in the relative peak areas of QC samples are excluded. Finally, the identification and relative quantification results of the metabolites are obtained. The PLS-DA model exhibited significant separation of clusters between the model and CAPE treatment groups (Figure 8A). Among the 1599 metabolites, 109 showed substantial differences in levels between the two groups. Specifically, CAPE-treated mice exhibited 19 metabolites with upregulated expression and 50 metabolites with downregulated expression compared to those in the control group (fold change > 1.5, p < 0.05) (Figure 8B). Furthermore, KEGG enrichment analysis revealed that the differentially abundant metabolites were significantly enriched in the purine metabolism and bile secretion pathways (Figure 8C). Adenosine is an endogenous purine nucleoside that binds to specific cell surface receptors to exert anti-inflammatory effects at high concentrations 31. Importantly, CAPE treatment of C. difficile-infected mice can increase the adenosine concentration in the gut (Figure 8D), which might be attributed to its inhibition of xanthine oxidase 32. Similarly, the levels of uric acid, the final product of purine degradation, were markedly decreased in mice after receiving CAPE (Figure 8D). An earlier study reported that increasing the gut adenosine concentration with an adenosine deaminase inhibitor prevents C. difficile toxin A-induced enteritis in mice 33. Therefore, the increase in gut adenosine induced by CAPE might play an important role in its therapeutic efficacy in CDI.

Alterations in gut metabolites in CAPE-treated mice with CDI.

(A) The score plots show orthogonal partial least square discriminant analysis (PLS-DA) results. (B) Volcano plot representing the upregulated or downregulated gut metabolites in the CAPE treatment group compared to those in the control group. (C) Enrichment bubble map of the top 20 metabolic pathways in the CAPE treatment and control groups. (D) The 26 potential biomarkers are displayed in a heatmap, along with hierarchical clustering analysis.

C. difficile predominantly utilizes sugars and amino acids for energy production through various metabolic pathways, such as fermentation 34. Notably, Stickland fermentation, a Clostridia-specific type of amino acid fermentation, is associated with C. difficile pathogenesis 34. D-proline is one of the preferred nutrients used by C. difficile in the Stickland reaction and is critical for its growth 35, 36. Interestingly, treatment of C. difficile-infected mice with CAPE significantly reduced gut D-proline levels (Figure 8D). In addition, the metabolomics data also revealed a dramatic increase in the level of melatonin (Figure 8D), which is a ubiquitous indolamine that was previously shown to decrease the rate of recurrent CDI 37. Consistently, quantification of the melatonin content in mouse fecal supernatants using a melatonin detection kit revealed elevated levels of melatonin in CAPE-treated mice (Figure S6). Moreover, correlation analysis revealed a positive association between Faecalibaculum abundance and melatonin levels (Figure S7). Taken together, these results indicate that alterations in gut metabolites caused by CAPE treatment might contribute to its therapeutic efficacy in CDI. We found that the metabolites of the gut microbiota changed after CAPE treatment. While detectable, the impact of gut microbiota modifications seems to be of secondary importance.

Oral administration of melatonin alleviates the symptoms of CDI

Next, we investigated whether any of the upregulated metabolites had a direct influence on TcdB. To this end, we preincubated recombinant TcdB with each of the selected metabolites prior to addition to the cell cultures. Interestingly, treatment with TcdB and 100 μg/mL melatonin significantly alleviated the cell death induced by this toxin (Figure 9A). This observation prompted us to investigate the direct therapeutic potential of melatonin in CDI. Therefore, C. difficile-infected mice were orally administered 200 mg/kg melatonin at 24 h intervals. The weight changes and the dry/wet ratio of the stool samples were monitored daily for three days. The results showed that melatonin-treated mice displayed remarkable alleviation of diarrhea, as manifested by notable improvements in weight recovery and a significant increase in the dry/wet weight ratio on the 2nd and 3rd days after C. difficile challenge (Figure 9B-C). Consistent with these observations, histopathological assessment of colonic tissue sections also revealed attenuated colon damage upon melatonin treatment (Figure 9D-E). Taken together, these data indicate that a high dose of exogenous melatonin could protect mice from CDI, However, the CAPE treatment group induced only minor alterations in gut microbiota-derived melatonin levels, which may not reach therapeutic relevance. While existing studies suggest that melatonin exhibits beneficial effects against CDI, the primary antimicrobial activity appears to be mediated directly by CAPE rather than through melatonin-associated mechanisms.

Oral administration of melatonin alleviates the pathology of CDI.

(A) The impact of upregulated metabolites on the cytotoxicity of TcdB. TcdB (0.2 ng/mL) was incubated with each of the metabolites (100 μg/mL) for 1 h at 37 °C before being added to the Vero cell culture. At 3 h after TcdB treatment, cell death was evaluated with a CCK-8 kit. (B) Body weight changes in melatonin-treated CDI mice (n=5). (C) The fecal dry/wet weight ratio in CDI mice after melatonin treatment (n=5). (D) Histopathological analysis of colon tissue from CDI mice treated with melatonin. (E) Histopathological scores of melatonin-treated CDI mice. * p <0.05; ** p <0.01 using one-way ANOVA.

Discussion

C. difficile has become the prevalent agent of nosocomial diarrhea and substantially threatens public health. In particular, the increase in morbidity and mortality caused by the prevalence of hypervirulent and drug-resistant strains has become a global concern 7. In 2019, C. difficile was listed by the US Centers for Disease Control and Prevention as an “urgent threat” that surpassed methicillin-resistant Staphylococcus aureus (MRSA) and ESBL-producing Enterobacterales 38. C. difficile is a commensal bacterium that colonizes the lower gastrointestinal tract 39. CDI typically occurs in susceptible individuals whose normal gut microbiota is disrupted by the administration of broad-spectrum antibiotics 39. Although CDI is considered an antibiotic-associated disease, the current treatment for CDI still relies on antimicrobial drugs, including metronidazole, fidaxomicin and vancomycin 9, 10. Despite antibiotic therapy, more than 35% of CDI patients are susceptible to recurrent infection, which is primarily attributable to the intrinsic effects of antimicrobial agents, which disrupt the microbiota 12. In addition, long-term administration of antibiotics promotes the selection of antibiotic-resistant subpopulations in the gut, thus increasing the risk of infections caused by antibiotic-resistant and opportunistic bacteria 40. As a result, the development of nonantibiotic therapeutics is needed for the treatment of CDI.

Indeed, numerous novel treatments for CDI are being investigated, among which the anti-virulence strategy represents one of the most promising and successful therapeutic approaches 22, 41. Unlike antibiotics, which target bacterial components that are critical for bacterial growth, anti-virulence drugs aim to disarm pathogenic bacteria by neutralizing their virulence factors 42, 43. Therefore, this strategy imposes less selective pressure that directs bacterial evolution of antimicrobial resistance 43. The disease progression of CDI is strictly dependent on large clostridial toxins (TcdA and TcdB), as strains lacking both toxins (i.e., A-B-) were completely avirulent and failed to induce any discernable disease symptoms in a hamster model of infection 44. Importantly, epidemic C. difficile strains with increased morbidity and mortality rates produce greater amounts of toxins 45. Based on these premises, TcdA and TcdB are considered ideal drug targets for CDI in the anti-virulence paradigm 46. Indeed, bezlotoxumab, a fully humanized monoclonal antibody that neutralizes TcdB, was approved by the FDA and used in CDI patients to reduce the rate of recurrence 47. This achievement has encouraged researchers to develop next-generation anti-virulence treatments. In particular, small molecule inhibitors of toxin production and/or activity have attracted increasing interest owing to their advantages in cost and the convenience of oral administration 41. The multidomain architecture and mechanisms of action of clostridial toxins have provided distinct options for blocking their toxicity. Indeed, previous efforts have identified diverse inhibitors that act on distinct steps of intoxication, such as the inhibition of toxin binding to cell receptors, the suppression of endosomal maturation, and the inactivation of GTD or CPD activities 23, 48, 49, 50, 51. Although which approach is the best remains unknown, we believe that an inhibitor that targets multiple steps of intoxication would be superior to those that act on a single step. In this study, we identified a series of caffeic acid derivatives that were potent inhibitors of TcdB via a phenotypic screen. Further mechanistic study revealed that CAPE, the phenethyl alcohol ester of caffeic acid, could block InsP6-induced autoproteolysis of TcdB and inactivate its GTD function (Figure 10). CAPE can also inhibit the growth of C. difficile and the production of spores. Moreover, CAPE treatment significantly protected against CDI in a mouse model of infection. Our results suggested that caffeic acid derivatives, particularly CAPE, are promising lead compounds for the development of targeted therapies against CDI. CAPE is a major component of propolis, a natural bee-produced substance that has been used as a remedy for thousands of years in Europe and Asia 52. As a result, further investigations on the therapeutic potential of propolis against CDI, either in combination with standard therapy or as a single therapy, would be of great interest.

A model for CAPE-mediated suppression of TcdB cytotoxicity and protection of mice from CDI.

The intoxication of cells by TcdB involves multistep mechanisms. Direct interaction of CAPE with full-length TcdB blocks InsP6-induced autoproteolysis, resulting in decreased production of active GTD. The binding between GTD and CAPE inhibits its glucosyltransferase activity, thus decreasing the glucosylation level of Rac1. CAPE treatment of CDI mice markedly restored the diversity and composition of the gut microbiota and induced changes in gut metabolites, which might have contributed to the therapeutic outcomes of CDI following CAPE treatment.

Antibiotic-induced gut microbiota dysbiosis is a major risk factor for CDI 53. Thus, reconstituting the gut microbiota is a promising strategy for CDI therapy. Fecal microbiota transplantation (FMT) is considered an effective CDI therapeutic strategy and is widely used in clinical practice 54. Despite the high cure rate and increasing popularity of FMT, patients remains limited acceptance. Another prominent advantage of anti-virulence drugs is that they allow maintenance of the normal microbiota, which is of critical importance for the treatment of CDI. Importantly, CAPE treatment of C. difficile-infected mice restored the structure and composition of the microbiota (Figure 10). In particular, the abundance of Bacteroides, a critical gut microbiota component that is required for resistance to C. difficile colonization 26, 55, 56, 57, was significantly elevated upon CAPE treatment. Interestingly, Bacteroides were also predominant in CDI patients administered RBX2660 (Rebyota), an FDA-approved fecal microbiota product 58, 59. The increased diversity of the gut microbiota might contribute to the therapeutic effects of CAPE in CDI.

In addition to the microbiota, gut metabolites also play crucial roles in the pathogenesis of CDI, either by promoting or suppressing disease progression 30. For example, bile acids and amino acids promote spore germination and pathogenesis, while secondary bile acid and short-chain fatty acids aid in colonization resistance and suppress spore germination 30. In addition to restoration of the gut microbiota, metabolomics analysis of mouse fecal samples revealed dramatic alterations in gut metabolites upon CAPE treatment (Figure 10). The differentially abundant metabolites were significantly enriched in the purine metabolism and bile metabolism pathways. Remarkably, the level of adenosine, an endogenous purine nucleoside that can regulate immune responses and inflammatory processes by binding to its receptors, such as its ability to inhibit the release of certain inflammatory mediators, was dramatically increased in CAPE-treated mice 60, 61. In contrast, the amount of uric acid, the final product of purine metabolism, was reduced. Previous studies showed that the increase in adenosine concentrations caused by the inhibition of adenosine deaminase via EHNA prevents C. difficile TcdA-mediated damage and inflammation 33. This protection might be a result of the anti-inflammatory effects of adenosine, which rely on the activation of its receptors 62. In addition, we observed a reduction in the amount of D-proline, which is one of the preferred amino acids utilized by C. difficile to generate energy through Stickland reactions 35. Melatonin is a ubiquitous indolamine that has broad biological activities 63. Notably, we detected a significantly increase in the amount of melatonin in the feces of mice that were administered CAPE. Importantly, an earlier study demonstrated a decreased risk of recurrent CDI in patients exposed to melatonin, possibly owing to its antibacterial and anti-inflammatory activities 37. In this study, we found that oral administration of exogenous melatonin alleviated the symptoms of CDI in a mouse model of infection. Although determining the exact mechanism will require further investigation, we propose that the melatonin-mediated inhibition of TcdB intoxication observed in this study might contribute to its therapeutic effect. Taken together, these results indicate that the alterations in gut metabolites caused by CAPE might also contribute to treatment outcomes for CDI patients.

Methods

Cell lines

All cell lines (HeLa, Caco-2, Vero, HEK293T and THP-1) were obtained from ATCC (Manassas, VA). HeLa, Caco-2, Vero and HEK293T cells were cultured in Dulbecco’s modified Eagle’s medium (DMEM) (Invitrogen, Carlsbad, CA) supplemented with 10% fetal bovine serum (FBS) (Invitrogen) and 1% (w/v) penicillin-streptomycin. THP-1 cells were cultured in Roswell Park Memorial Institute (RPMI) 1640 medium (Invitrogen) supplemented with 10% FBS and 1% (w/v) penicillin-streptomycin. All cell lines were cultured at 37 °C in a 5% CO2 atmosphere.

Bacterial strains and reagents

C. difficile BAA-1870 (RT027/ST1) was kindly provided by Aiwu Wu (Guangzhou Medical University) and cultured in brain-heart infusion broth (BHI) supplemented with 5 g/L yeast extract and 0.1% (w/v) L-cysteine (BHIS). The compounds were identified from two main sources: FDA-approved drugs obtained from ApexBio Technology (Houston, TX, USA) and small molecule natural compounds obtained from Chengdu Ruifensi Biotechnology Co., Ltd. (Chengdu, China). LysoTracker Red DND-99, kanamycin, and cefoperazone sodium were purchased from Dalian Meilun Biotechnology Co., Ltd. (Dalian, China).

Mice

Male C57BL/6 mice aged 6-8 weeks were purchased from Liaoning Changsheng Biotechnology Co., Ltd., and housed in a specific pathogen-free (SPF) facility. The mice lived in a controlled environment with a temperature of 23±2 °C and a humidity of 55±10 °C. Prior to infection, the mice were acclimated to the environment for a minimum of one week. All mouse experiments were approved by the Jilin University Institutional Animal Care Committee, Jilin University and strictly conducted according to the guidelines of this committee (SY202402301).

Protein expression and purification

The expression and purification of full-length TcdB were conducted following the method described by Yang et al. 64. In brief, Bacillus megaterium harboring pHis1522-TcdB was inoculated into LB media supplemented with a final concentration of 10 μg/mL tetracycline. Once the OD600nm reached approximately 0.3, xylose (0.5% w/v) was added to induce protein expression at 37 °C for 16 h. Then, 100 mL of the bacterial cells was harvested by centrifugation, resuspended in 5 mL of lysis buffer containing 20 mM imidazole, 300 mM NaCl, 20 mM NaH2PO4, 500 μM EDTA, and protease inhibitor cocktail (adjusted to pH 8.0), and subjected to sonication. The cell lysate was then centrifuged at 13,400 × g for 20 min to remove cell debris. The TcdB protein was purified using nickel column affinity chromatography. The purity of the protein was confirmed by SDS-PAGE, and the protein was subsequently dialyzed against buffer containing 10% glycerol, 20 mM Tris (pH 7.5), and 150 mM NaCl. After the protein concentration was determined, the purified TcdB protein was aliquoted and stored at −80 °C for further use.

For the purification of GTD, amplified DNA products were inserted into pET28a (Novagen), and the Escherichia coli (E. coli) strain BL21 (DE3) was used for the expression and purification of GTD. Transformed E. coli pET28a-GTD was cultured in Luria–Bertani (LB) medium supplemented with 10 μg/mL kanamycin. Once the OD600 reached 0.6, the expression of proteins was induced with 0.2 mM isopropyl-β-d-thiogalactoside (IPTG) overnight at 18 °C. The bacterial cells were then harvested by centrifugation and subjected to sonication. The cell lysate was then centrifuged at 13,400 × g for 20 min to remove cell debris. GTD protein was purified using nickel affinity chromatography and dialyzed in buffer containing 10% glycerol, 20 mM Tris (pH 7.5), and 150 mM NaCl. The GTD mutants were introduced into pET28a-GTD using the Quick Change Site-Directed Mutagenesis Kit (Stratagene, San Diego, CA, USA) with the primers listed in Table S2. The subsequent purification was performed using the same procedure as described for wild-type GTD.

Reverse transcription-quantitative PCR

Total RNA was extracted by TRIzol and then reverse-transcribed into complementary DNA by a NovoScript® Plus All-in-one 1st Strand cDNA Synthesis SuperMix (E047; Novoprotein, China). mRNA expression of tcdB was measured by quantitative PCR with NovoScript®SYBR qPCR SuperMix Plus (E096; Novoprotein, China) using the primers shown in Table S3. 16S rRNA was used as internal control. The relative mRNA expression levels were calculated by the delta-delta CT method.

Growth curve assays

Growth curve assay was performed to estimate the influence of CAPE on the growth of C. difficile BAA-1870. The overnight culture of BAA-1870 was transferred to fresh BHIS medium and incubated at 37°C in an anaerobic atmosphere. The culture was supplemented with different concentrations of CAPE ranging from 0 to 64 μg/mL and the optical density at 600 nm was monitored at 4 h intervals to assess growth of BAA-1870.

Inhibition of spore outgrowth

The impact of CAPE on the sporulation of C. difficile BAA-1870 was assessed using the protocol described by Joshua et al 65. In brief, BAA-1870 cultures were treated with 8 or 32 μg/mL of CAPE for 48 h. The treated samples were then harvested, subjected to a double wash with PBS, and incubated in a 60°C water bath for 30 min. Subsequently, 20 μL of each spore suspension was plated onto BHIS agar supplemented with 0.1% taurine to facilitate germination. These plates were incubated anaerobically at 37°C for 5 d to allow for spore germination. Following this period, the number of resulting colonies on the plates was enumerated.

Cell rounding assay

Cell rounding assay was performed according to Kristina et al, with appropriate adjustments 48. Vero or HeLa cells were seeded in 96-well sterile plates using serum-free DMEM overnight, and the compounds were preincubated with TcdB for 1 h. After incubation, the TcdB-compound complex was added to cells for 1 h at 37 °C. The cell morphology was visualized under a light microscope.

LysoTracker assay

The pH of endosomes was determined by LysoTracker Red DND-99 49. Caco-2 or THP-1 cells in complete DMEM were seeded in 96-well sterile plates overnight. The media was changed to serum-free DMEM for 1 h, and CAPE was preincubated with TcdB for 1 h, followed by addition to the cells and incubation for 2 h. Baf-A1 served as a control. After incubation, LysoTracker and Hoechst dyes were added and the cells were incubated for 1 h. Excess dye was removed by changing the media, and alterations in lysosomal red intensity were documented using a fluorescence microscope.

In vitro TcdB autoprocessing assay

The self-cleavage ability of TcdB primarily depends on its cysteine protease activity 48. CAPE (1-16 μg/mL) was preincubated with TcdB for 1 h, followed by the addition of InsP6 and incubation at 37 °C for 6 h. Subsequently, loading buffer was added for SDS-PAGE analysis, and the results were visualized using Coomassie Brilliant Blue staining.

Rac1 glucosylation assay

Rac1 glucosylation assay was performed following the method described by John et al 50. HEK293T cells were seeded into a 6-well plate. The next day, TcdB and different concentrations of CAPE (4-32 μg/mL) were added to cells. After 2 h, the cells were harvested, lysed using lysis buffer, and subjected to centrifugation at 13,400 × g for 10 min to collect the supernatant. The collected supernatant was then mixed with loading buffer and subjected to SDS-PAGE followed by western blot analysis. The glycosylation level of Rac1 was evaluated by determining the levels of non-glycosylated Rac1 (BD Biosciences, Mississauga, ON) and total Rac1 (Abcam).

UDP-GloTM UDP-glucose hydrolase assay

GTD activity was assessed according to the manufacturer’s instructions 50. A total volume of 16 μL, consisting of 100 nM GTD, various concentrations of CAPE and glucosylation buffer, was prepared. Four microliters of UDP-glucose at a final concentration of 25 μM was added to the mixture, which was then allowed to react at room temperature for 15 min. To stop the reaction, 10 μL of the reaction mixture was removed and placed in a white polystyrene 96-well plate containing 10 μL of UDP detection reagent. The reaction took place at room temperature for 1 h. The luminescence value was measured at a 750 ms integral, and the GTD activity was determined based on these measurements.

Circular dichroism (CD) spectroscopy

CD spectra of TcdB, TcdB-CAPE, GTD, and GTD-CAPE were obtained using a CD spectrophotometer (MOS-500; Bio-Logic, France). TcdB or GTD at a concentration of 0.2 mg/mL was incubated with 32 μg/mL CAPE. Changes in the secondary structure of TcdB or GTD were assessed at room temperature using a quartz cuvette with an optical path length of 1 mm. The scan wavelength range was set from 190 to 250 nm with a resolution of 0.2 nm and a bandwidth of 1 nm. The secondary structures of TcdB, TcdB-CAPE, GTD, and GTD-CAPE were analyzed using the BeStSel web server.

Molecular docking

CAPE was obtained as a ligand from PubChem, and its topology was generated using AmeberTools. The crystal structure of GTD was retrieved from the Protein Data Bank (PDB) and processed using AutoDockTools. Docking calculations were performed using AutodockVina, and molecular dynamics simulations were carried out using Gromacs 2020.6. The RMSD of the receptor and ligand was monitored to assess conformational changes during the binding process. The interaction energy was analyzed using the Poisson-Boltzmann surface area (MM-PBSA) method, which is a molecular mechanics approach.

Flow cytometry

Flow cytometry was used to quantify the cellular uptake of TcdB via endocytosis. To facilitate detection, full-length TcdB was labeled with Alexa Fluor®488 NHS Ester (Yeasen Biotechnology Co., Ltd., Shanghai, China). Prior to cell treatment, Alexa Fluor 488-TcdB was preincubated with CAPE at a concentration of 32 μg/mL for 30 min. The treated solution was then added to Caco-2 cells and incubated for 1 h at 37 °C. Following three washes with phosphate-buffered saline (PBS), the Caco-2 cells were collected, and the fluorescence signal was measured using flow cytometry.

Cytotoxicity assay

Vero or THP-1 cells were plated into 96-well plates. The next day, different concentrations of CAPE were prepared by diluting them in DMEM and adding them to the cells. After incubation at 37 °C for 3 h, the viability of cells was assessed using a Cell Counting Kit-8 (CCK-8) (Beyotime Biotechnology Co., Shanghai, China) according to the manufacturer’s instructions.

Biacore analysis

The affinity analysis of TcdB or GTD with CAPE was conducted using a Biacore T100 device (Biacore Inc., Uppsala, Sweden) according to the manufacturer’s protocol. A CM5 chip (General Electric Company, GE) was utilized to immobilize the TcdB or GTD proteins. Various concentrations of CAPE were injected, and all steps were performed in running buffer consisting of 5% DMSO in PBS-P20 (GE). Evaluation version 3.1 software provided by Biacore and the 1:1 Langmuir combined model were employed for data analysis.

Mouse infection model

For the acute toxicity model, after pretreatment of TcdB with CAPE (2 or 4 μg/mL) for 1 h, the mice were intraperitoneally injected with a complex of TcdB-solvent or TcdB-CAPE (30 ng of TcdB per mouse). The survival of mice was observed for 72 h.

For the CDI model, mice were given cefoperazone (0.5 g/L) in their drinking water for 5 days (from day −7 to day −2), followed by switching to sterilized water for 2 days. On day 0, the mice were challenged with 1 × 108 colony-forming units (CFUs) of C. difficile BAA-1870 (Ribotype 027) by oral gavage. For the CAPE or melatonin treatment group, mice were treated with CAPE (30 mg/kg by oral gavage) or melatonin (200 mg/kg by oral gavage) every day after C. difficile challenge. The weight loss and fecal state of the mice were monitored daily. Fecal samples were collected on days 0-3 to evaluate the stool water content as a marker of diarrhea. Additionally, fecal samples on day 3 were collected to assess C. difficile colonization. Colonic tissues were collected on day 3 post infection, stained with hematoxylin and eosin (H&E) and scored as previously described 66.

Fecal samples were collected on day 3 post infection, frozen in liquid nitrogen and transferred to the laboratory with dry ice for 16S rRNA sequencing and untargeted metabolomics analysis (Novogene Bioinformatics Technology Co., Ltd., Tianjin, China). The melatonin level in mouse stool supernatant was determined using the Mouse MT (Melatonin) ELISA Kit (Sangon Biotechnology Co., Ltd., Shanghai, China).

Statistical analysis

Western blot densitometry and LysoTracker fluorescence intensity were measured using ImageJ software (NIH, Bethesda, MD, USA). Statistical analysis was performed using GraphPad Prism 9.0 (GraphPad Software, USA). Unpaired two-tailed Student’s t tests were used for comparing two groups, and one-way analysis of variance (ANOVA) was used for analyzing data from more than three groups. The statistical significance was determined using the log-rank (Mantel–Cox) test (survival rates) and Mann–Whitney U test (tissue bacterial load) in vivo. All the data are expressed as the mean ± standard deviation (SD). * indicates p <0.05 and ** indicates p <0.01.